---
title: "COVID-19 Mapper"
format: html
---

This document attempts to recreate the work in the paper "Topological data analysis model for the spread of coronavirus" by Chen and volilc (2021).
The aim of this paper is to understand and simplify the United States COVID-19 data cloud whilst preserving the underlying structure and features.

Todo: summarise each individual component (bar chart of states/counties, cases), create annotated image of the components

First I need to clean and preprocess the data in accordance to their method:
* days = number of days since 22/1/20 (british format)
* select longitude, lattitude, cummulative confirmed cases, days since
* Create two subsets of the data: unfiltered, and filtered (no Los Angeles, Illinois or New York)
  + Further two subsets for the temporal ranges to 19/6/20 and 24/7/20
  + For now, I will just use the first boundary, as to simplify the document

```{r}
#| label: data-setup
#| output: false

library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(MapperAlgo)
library(igraph)

raw_data <- read.csv(here::here("data/covid_19_jhu/RAW_us_confirmed_cases.csv"))

start_date <- as.Date('22/1/20', format = "%d/%m/%y")
end_date1 <- as.Date('19/6/20', format = "%d/%m/%y")
end_date2 <- as.Date('24/7/20', format = "%d/%m/%y")

# create long date format
# quick data formatting
rename_cols <- c(
    state = "Province_State",
    county = "Admin2",
    uid = "UID",
    fips = "FIPS",
    country_region = "Country_Region",
    latitude = "Lat",
    longitude = "Long_",
    combined_key = "Combined_Key"
)
factor_cols <- c("state", "county", "iso2", "iso3", "country_region")
long_data <- raw_data |>
    pivot_longer(
        cols = starts_with("X"),
        names_to = "date",
        values_to = "cases"
    ) |>
    mutate(
        date = gsub("^X", "", date),
        date = gsub("\\.", "/", date),
        date = mdy(date),
        days = as.integer(difftime(date, start_date, units = "days")),
        index = row_number()
    ) |>
    filter(
        date >= start_date
    ) |>
    rename(all_of(rename_cols)) |>
    mutate(across(all_of(factor_cols), as.factor))

# create subsets for different ranges
# subset2 <- long_data |>
#     filter(date <= end_date2)

geographically_filter_data <- function(data) {
    filtered <- data |>
        filter_out(when_any(
            county %in% c("Los Angeles"),
            state %in% c("Illinois", "New York")
        ))
    return(filtered)
}

subset1 <- long_data |>
    filter(date <= end_date1)

# scale so the sum of the squared column values is 1 (L2 norm)
# used when direction is more important than magnitude
# outliers kept as this complements DBSCAN
filter_cols <- c("latitude", "longitude", "cases", "days")
subset1 <- subset1 |>
    mutate(across(
        all_of(filter_cols),
        function(col) {
            norm_val <- sqrt(sum(col^2, na.rm = TRUE))
            if (norm_val == 0) return(col) # stop division error
            return(col / norm_val)
        },
        .names = "{.col}_norm"
    ))

# keep the unfiltered norm values, just removing the known outliers
filt_subset1 <- geographically_filter_data(subset1)
```

# Mapper Setup

Now we can try the Mapper algorithm used:
* No filter function per se, just the identity mapping of the data
* Euclidean distance metric
* 10 intervals, 8% overlap
* DBSCAN clustering

There filter function is is just the identity map (a 4-dimensional function that is just the data points).
With such few dimensions, and the fact they are exploring for statistical findings, there was no need to project the data.
An 8% overlap was determined as it "empirically" appears to give the most informative Mapper graphs.
DBSCAN was simply chosen

The colours for the graphs are based on the data's order, which is by the geographical information, so nearby states are coloured similary.

```{r}
#| label: mapper setup

base_dir <- file.path("examples", "covid_mapper_files")
subset1_mapper_file <- "subset1_mapper.json"
filt_subset1_mapper_file <- "filt_subset1_mapper.json"
subset1_mapper_data <- subset1 |> select(ends_with("_norm"))
filt_subset1_mapper_data <- filt_subset1 |> select(ends_with("_norm"))
```

```{r}
#| label: create mappers

subset1_mapper_time <- system.time({
    subset1_mapper <- MapperAlgo(
        subset1_mapper_data,
        filter_values = subset1_mapper_data,
        percent_overlap = 8,
        intervals = 10,
        methods = "dbscan",
        method_params = list(eps = 0.5, minPts = 3), # same config as KeplerMapper
        num_cores = 18 # quicker performance
    )
})
message("Unfiltered subset1 Mapper time:", subset1_mapper_time)

export_data <- list(
  adjacency = subset1_mapper$adjacency,
  num_vertices = subset1_mapper$num_vertices,
  level_of_vertex = subset1_mapper$level_of_vertex,
  points_in_vertex = subset1_mapper$points_in_vertex,
  original_data = subset1_mapper_data
)
write(toJSON(export_data, auto_unbox = TRUE), file.path(base_dir, subset1_mapper_file))
message("Saved mapper as JSON to ", file.path(base_dir, subset1_mapper_file))

filt_subset1_mapper_time <- system.time({
    filt_subset1_mapper <- MapperAlgo(
        filt_subset1_mapper_data,
        filter_values = filt_subset1_mapper_data,
        percent_overlap = 8,
        intervals = 10,
        methods = "dbscan",
        method_params = list(eps = 0.5, minPts = 3), # same config as KeplerMapper
        num_cores = 18 # quicker performance
    )
})
message("Filtered Mapper time:", filt_subset1_mapper_time)

export_data <- list(
  adjacency = filt_subset1_mapper$adjacency,
  num_vertices = filt_subset1_mapper$num_vertices,
  level_of_vertex = filt_subset1_mapper$level_of_vertex,
  points_in_vertex = filt_subset1_mapper$points_in_vertex,
  original_data = filt_subset1_mapper_data
)
write(toJSON(export_data, auto_unbox = TRUE), file.path(base_dir, filt_subset1_mapper_file))
message("Saved mapper as JSON to ", file.path(base_dir, filt_subset1_mapper_file))
```

```{r}
#| label: load mappers
subset1_mapper <- fromJSON(file.path(base_dir, subset1_mapper_file))
filt_subset1_mapper <- fromJSON(file.path(base_dir, filt_subset1_mapper_file))
```

# Mapper Exploration
```{r}
#| label: mapper_visualisation

subset1_plot <- MapperPlotter(
    subset1_mapper,
    original_data = subset1_mapper_data,
    label = subset1$latitude,
    avg = TRUE,
    use_embedding = FALSE
)
subset1_plot

subset1_graph <- graph_from_adjacency_matrix(subset1_mapper$adjacency, mode = "undirected")
V(subset1_graph)$

filt_subset1_plot <- MapperPlotter(
    filt_subset1_mapper,
    original_data = filt_subset1_mapper_data,
    label = filt_subset1$latitude,
    avg = TRUE,
    use_embedding = FALSE
)
filt_subset1_plot
```


```{r}
#| label: node data
source("utils.R")
sourceDir("mapper", exclude = "test.R")

subset1_mapper_stats <- summarise_graph(subset1_mapper, subset1_mapper_data)
print(subset1_mapper_stats)

original_nrow <- nrow(subset1)

subset1_nodes <- get_node_mapping(subset1_mapper, subset1)
subset1 <- subset1 |>
    # make a new row for every node assignment
    left_join(subset1_nodes, by = "index") |>
    unnest_longer(node)

new_nrow <- nrow(subset1)
diff <- new_nrow - original_nrow
if (diff > 0) {
    message("Subset1 increased by ", diff, " rows")
} else {
    message("No dimensional change for subset 1")
}

original_nrow <- nrow(filt_subset1)

filt_subset1_nodes <- get_node_mapping(filt_subset1_mapper, filt_subset1)
filt_subset1 <- filt_subset1 |>
    left_join(filt_subset1_nodes, by = "index") |>
    unnest_longer(node)

new_nrow <- nrow(filt_subset1)
diff <- new_nrow - original_nrow
if (diff > 0) {
    message("Filtered subset1 increased by ", diff, " rows")
} else {
    message("No dimensional change for filtered subset1")
}
```

I will explore the clusters.

```{r}
#| label: get_clusters

subset1_clusters <- find_isolated_components(subset1_mapper, subset1)
# check to see if there are nodes in multiple components (sanity check)
subset1_clusters |>
   group_by(node) |>
   summarise(n_comps = n()) |>
   filter(n_comps > 1)

subset1 <- subset1 |>
    left_join(subset1_clusters, by = "node")

filt_subset1_clusters <- find_isolated_components(filt_subset1_mapper, filt_subset1)
filt_subset1_clusters |>
   group_by(node) |>
   summarise(n_comps = n()) |>
   filter(n_comps > 1)

filt_subset1 <- filt_subset1 |>
    left_join(filt_subset1_clusters, by = "node")
```

### Component 1

Through investigation, component 1 is the American Samoa, located near the islands Fiji and Samoa.
It consists of 10 nodes, each with 16 rows of data.
This component has 0 cases of COVID-19, which could be a systematic error (no measurements taken).
```{r}
#| label: unfiltered cluster 1

subset1_comp1 <- subset1 |>
    filter(component == 1)
summary(subset1_comp1)
table(subset1_comp1$state, useNA = "no")
table(subset1_comp1$county)
```

### Component 2

Component 2 is Hawaii.
```{r}
#| label: unfiltered cluster 2
subset1_comp2 <- subset1 |>
    filter(component == 2)

summary(subset1_comp2)
table(subset1_comp2$state) # only Hawaii
message("Time range: ", min(subset1_comp2$date), " to ", max(subset1_comp2$date))
```

### Component 3

Component 3 is more western Alaska, consisting of 4160 data points and 40 nodes, with consistent large peaks and small ones also.
```{r}
#| label: unfiltered cluster 3

subset1_comp3 <- subset1 |>
    filter(component == 3)

summary(subset1_comp3)
table(subset1_comp3$node)
length(unique(subset1_comp3$node))
ggplot(subset1_comp3, aes(node)) + geom_histogram()
```

### Component 4
Component 4 is the main trunk of data, which I will now explore.

```{r}
#| label: main_trunk

unfilt_trunk_data <- unfilt_mapper_raw |>
    filter(component == 4)

unfilt_trunk_data |>
    group_by(state) |>
    summarise(n = n()) |>
    arrange(desc(n))
```

### Component 5

Component 5 is Eastern Alaska
```{r}
#| label: unfiltered cluster 5
subset1_comp5 <- subset1 |> filter(component == 5)

summary(subset1_comp5)
table(subset1_comp5$node)
```

### Component 6

This component consists of all of Puerto Rico
```{r}
#| label: unfiltered cluster 6

subset1_comp6 <- subset1 |> filter(component == 6)
summary(subset1_comp6)
table(subset1_comp6$node)
```

### Component 7


```{r}
#| label: unfiltered cluster 7

subset1_comp7 <- subset1 |> filter(component == 7)
summary(subset1_comp7)
table(subset1_comp7$node)
```

### Component 8

Component 8 is Guam and the Northern Mariana Islands in the North-Western Pacific
```{r}
#| label: unfiltered cluster 8

subset1_comp8 <- subset1 |> filter(component == 8)
summary(subset1_comp8)
table(subset1_comp8$node)
```

### Component 9

Component 9 is just Cook and Illinois from 04/05/2020 to 15/05/2020, just 11 days.
This is highlighted as its own component as it consists of a large amount of cases, more than 250 times the data's mean 
```{r}
#| label: unfiltered cluster 9

subset1_comp9 <- subset1 |> filter(component == 9)
summary(subset1_comp9)
table(subset1_comp9$node)
```

### Component 10

This is more Cook, Illinois in the 17,000 range of cases from 16/05/2020 to 19/16/2020
```{r}
#| label: unfiltered cluster 10

subset1_comp10 <- subset1 |> filter(component == 10)
summary(subset1_comp10)
table(subset1_comp10$node)
```

### Component 11

This is Harris, Texas, from 09/06/2020 to 19/06/2020
```{r}
#| label: unfiltered cluster 11

subset1_comp11 <- subset1 |> filter(component == 11)
summary(subset1_comp11)
table(subset1_comp11$node)
```

### Component 12

This is a section of Los Angeles, California, from 05/06/2020 to 17/06/2020 recording between 61000 and 78000 cases in this component.
```{r}
#| label: unfiltered cluster 12

subset1_comp12 <- subset1 |> filter(component == 12)
summary(subset1_comp12)
table(subset1_comp12$node)
```

### Component 13

This is the final LA section of nearly 80,000 cases. It is a single node of 2 data points.
```{r}
#| label: unfiltered cluster 13

subset1_comp13 <- subset1 |> filter(component == 13)
summary(subset1_comp13)
table(subset1_comp13$node)
```

# Smaller Experimentation

I will try explore a smaller subset of data between 22/1/2020 to 5/4/2020 (the median)


```{r}
small_data <- subset1 |>
    filter(date < as.Date("5/4/20", format = "%d/%m/%y"))

small_scaled <- small_data |>
    mutate(across(all_of(filter_cols), scale(.data, center = FALSE)), .names = "{.col}_zscale")
```