---
title: "Open Play: Data Understanding"
author: "Daniel Agar"
format:
  pdf: default
  html:
    toc: true
  docx: default
abstract: "This document contains the initial exploration of the Open Play dataset, but focuses on getting an analysis-ready intake dataset."
---

This document explores the clean version of the data set Open Play.
This is to help me get a better understanding of the data I will be working with for my MSc dissertation.
The cleaning process consisted of the following:
- Steam:
  -Removed idle games or non games (SteamVR, soundpad, etc.),
  - Added approximate session times
  - Scaled sessions over an hour
  - Boolean indicator for new sessions
  - Added number of hours sessions lasted
  - Calculated metrics for each session/game id
  - Created quality metrics (separate file containing information about the scaled instances)
- Xbox:
  - Merge adjacent sessions that are within a threshold (60 seconds)
  - Remove non-game sessions (shopping/videos)
  - Remove suspiciously long sessions (more than 6 hours)
  - Merges adjacent sessions by splitting into lots of atomic sessions and merging
- Nintendo:
  - Filter duplicates
  - Merge adjacent sessions with a minute tolerance
- Android: simple renaming
- iOS: simple renaming
- Excludes entries where total gaming for a day is more than 16 hours
- intake:
  - Self-reported hours > 50, NA
  - Remove duplicates of linked_platforms
- daily:
  - Remove daily duplicates of survey
  - Recode erroneous wave codings by their start date
  - Ignore "false starts" of game data (outside date range)
- bikweekly:
  - Remove daily duplicates of survey
  - Recode erroneous wave codings by their start date
  - Ignore "false starts" of game data (outside date range)
- timeuse: assign quality flags to diaries
  - No mention of sleep or personal care
  - No sleep
  - More than 90 minutes unacounted for
  - Less than 5 activites logged during a day
- simon:
  - Only include non-pilot-study participants (weren't in initial pre-trial) who qualified at intake

```{r} 
#| label: libraries
#| include: false
#| code-summary: "Show code (Load required R packages)"
library(tidyverse)
library(data.table)
library(dtplyr)
library(naniar)
library(gtsummary)
library(GGally)
library(patchwork)
library(VIM)
library(mice)
library(ggmice)
library(car)
library(skimr)

theme_set(theme_minimal())
source("examples/open-play/helpers.R")
```

I will fetch the cleaned datasets:
```{r}
#| label: data setup
#| include: false
#| cache: true
android <- read_csv("data/open-play/clean/android.csv.gz")
metadata <- read_csv("data/open-play/clean/game_metadata.csv.gz")
ios <- read_csv("data/open-play/clean/ios.csv.gz")
nintendo <- read_csv("data/open-play/clean/nintendo.csv.gz")
simon <- read_csv("data/open-play/clean/simon.csv.gz")
steam <- read_csv("data/open-play/clean/steam.csv.gz")
steam_visibility <- read_csv("data/open-play/raw/telemetry_steam_account_linking_raw.csv.gz")
biweekly <- read_csv("data/open-play/clean/survey_biweekly.csv.gz")
daily <- read_csv("data/open-play/clean/survey_daily.csv.gz", guess_max = 10000)
intake <- read_csv("data/open-play/clean/survey_intake.csv.gz", guess_max = 34000)
time_use <- read_csv("data/open-play/clean/timeuse.csv.gz") |>
  filter(quality_flag == "Good quality")
xbox <- read_csv("data/open-play/clean/xbox.csv.gz")

xbox_raw <- read_csv("data/open-play/raw/telemetry_xbox_raw.csv.gz")
nintendo_raw <- read_csv("data/open-play/raw/telemetry_nintendo_raw.csv.gz")
daily_raw <- read_csv("data/open-play/raw/survey_daily_raw.csv.gz")
biweekly_raw <- read_csv("data/open-play/raw/survey_biweekly_raw.csv.gz", guess_max = 5000)
time_use_raw <- read_csv("data/open-play/raw/timeuse_raw.csv.gz")
```

```{r}
#| label: data-prep
#| cache: true
# clean most likely to use columns
steam_visibility <- steam_visibility |>
  filter(platform == "steam") |>
  mutate(
    timestamp = as_datetime(timestamp, tz = "UTC"),
    is_public = event == "steamDataPublic"
  ) |>
  select(pid, timestamp, event, is_public)

# import raw data for calculating how many rows were excluded
xbox_valid_sessions <- xbox_raw |>
  merge_adjacent_sessions(tol_sec = 60) |>
  # filter non-foreground games
  filter(
    duration >= 1,
    pid %in% intake$pid,
    title_placement == "Full", # in foreground
    !genres %in% c("Shopping", "Video", "Entertainment", "Utilities & tools"),
    tolower(title_id) != "e1924c10-9c91-4652-81b9-8ca0670e77a7" # home screen
  )

nintendo_valid_sessions <- nintendo_raw |>
  distinct() |>
  filter(
    pid %in% intake$pid,
    duration >= 1
  ) |>
  merge_adjacent_sessions(tol_sec = 60)
```

Now I shall select the qualified participants, as I know I will have more data to analyse.
These participants are known to have completed at least one (biweekly) survey.

```{r}
#| label: aggregate-session-hourly-daily
#| cache: true
# merge with the local time zone offsets from intake
# Note: We keep country and local_timezone for DST-aware conversion
tz_map <- intake |>
  mutate(
    pid,
    country,
    local_timezone,
    .keep = "none"
  ) |>
  distinct(pid, .keep_all = TRUE)

# aggregate data at each level of granularity (session, hourly, daily)
# session level (Nintendo and Xbox)
session_telemetry <- bind_rows(
    xbox |> mutate(platform = "Xbox"),
    nintendo |> mutate(platform = "Nintendo")
  ) |>
  left_join(tz_map, by = "pid") |>
  filter(!is.na(local_timezone)) |>
  mutate(
    # Calculate DST-aware offset for each timestamp
    offset_start = get_dst_offset(session_start, country, local_timezone),
    offset_end = get_dst_offset(session_end, country, local_timezone),
    # Add offset to get local time values (keeping UTC label for compatibility)
    start_local = session_start + offset_start,
    end_local = session_end + offset_end,
    duration_min = as.numeric(difftime(
      session_end,
      session_start,
      units = "mins"
    ))
  ) |>
  filter(
    !is.na(session_start),
    !is.na(session_end),
    session_end > session_start,
    duration_min >= 1
  )

# Hourly (Nintendo and Xbox)
# expand sessions into local-hour bins, compute overlap minutes, and add UTC hour
hourly_from_sessions <- session_telemetry |>
  filter(!is.na(start_local), !is.na(end_local)) |>
  mutate(
    h0_local = floor_date(start_local, "hour"),
    h1_local = floor_date(end_local - seconds(1), "hour"),
    n_hours = as.integer(difftime(h1_local, h0_local, units = "hours")) + 1
  ) |>
  filter(!is.na(n_hours), n_hours > 0) |>
  tidyr::uncount(n_hours, .remove = FALSE, .id = "k") |>
  mutate(
    hour_start_local = h0_local + hours(k - 1),
    minutes = pmax(
      0,
      as.numeric(difftime(
        pmin(end_local, hour_start_local + hours(1)),
        pmax(start_local, hour_start_local),
        units = "mins"
      ))
    ),
    # Convert back to UTC (this preserves the instant, just changes label)
    hour_start_utc = with_tz(hour_start_local, tzone = "UTC")
  ) |>
  select(pid, platform, title_id, hour_start_local, hour_start_utc, minutes) |>
  group_by(pid, platform, title_id, hour_start_local, hour_start_utc) |>
  summarise(minutes = sum(minutes, na.rm = TRUE), .groups = "drop")

hourly_from_steam <- steam |>
  mutate(
    datetime_hour_start = ymd_hms(
      paste0(date, " ", sprintf("%02d", hour), ":00:00"),
      tz = "UTC"
    )
  ) |>
  select(pid, title_id, datetime_hour_start, minutes) |>
  left_join(tz_map, by = "pid") |>
  filter(!is.na(local_timezone)) |>
  mutate(
    platform = "Steam",
    hour_start_utc = datetime_hour_start,
    offset = get_dst_offset(datetime_hour_start, country, local_timezone),
    # Note: These are local time values with UTC labels due to R's POSIXct limitation
    hour_start_local = datetime_hour_start + offset
  ) |>
  select(pid, platform, title_id, hour_start_local, hour_start_utc, minutes) |>
  group_by(pid, platform, title_id, hour_start_local, hour_start_utc) |>
  summarise(minutes = sum(minutes, na.rm = TRUE), .groups = "drop")

hourly_telemetry <- bind_rows(hourly_from_sessions, hourly_from_steam)

# Daily (Nintendo + Xbox + Steam; collapse hourly to days)
daily_telemetry <- hourly_telemetry |>
  mutate(
    day_local = as.Date(hour_start_local),
  ) |>
  group_by(pid, platform, day_local) |>
  summarise(minutes = sum(minutes, na.rm = TRUE), .groups = "drop") |>
  bind_rows(ios) |>
  bind_rows(android)

# Weekly
weekly_all <- daily_telemetry |>
  mutate(
    week = floor_date(day_local, "week")
  ) |>
  group_by(pid, platform, week) |>
  summarise(minutes = sum(minutes, na.rm = TRUE), .groups = "drop")

telemetry_spans <- daily_telemetry |>
  group_by(pid, platform) |>
  summarise(
    telemetry_start = min(day_local, na.rm = TRUE),
    telemetry_end = max(day_local, na.rm = TRUE) + hours(1), # end of last hour bin
    week = floor_date(telemetry_end, "week"),
    n_weeks = as.integer(difftime(
      telemetry_end,
      telemetry_start,
      units = "weeks"
    )) +
      1,
    .groups = "drop"
  )
```

```{r}
#| label: participant-summary-stats
#| cache: true

# participants
n_baseline <- nrow(intake[intake$qualified, ])
# participants in surveys
n_participants <- sum(unique(c(daily$pid, biweekly$pid)) %in% intake$pid)
n_participants_biweekly <- sum(unique(c(biweekly$pid)) %in% intake$pid)
n_participants_daily <- sum(unique(c(daily$pid)) %in% intake$pid)

participant_age <- (function(x) {
  sprintf(
    "%s years old (10th percentile: %s, 90th percentile: %s)",
    print_num(median(x), .1),
    print_num(quantile(x, 0.10), .1),
    print_num(quantile(x, 0.90), .1)
  )
})(na.omit(intake$age[intake$qualified]))

n_daily <- nrow(daily)
n_biweekly <- nrow(biweekly)
n_sessions <- (nrow(nintendo) + nrow(xbox) + nrow(steam))
n_hours <- (
  sum(nintendo$duration, na.rm = TRUE) +
  sum(xbox$duration, na.rm = TRUE) +
  sum(steam$minutes, na.rm = TRUE) +
  sum(ios$minutes, na.rm = TRUE) +
  sum(android$minutes, na.rm = TRUE)) /
  60

n_platforms <- table(
  str_count(
    intake$linked_platforms[!is.na(intake$linked_platforms)],
    ","
  ) +
    1
)

n_hours <- sum(daily_telemetry$minutes / 60)
n_users <- length(unique(daily_telemetry$pid))
n_games <- length(unique(hourly_telemetry$title_id))
n_months <- round(max(telemetry_spans$n_weeks) / 4.2)
n_surveys <- nrow(daily) + nrow(biweekly)
n_simon <- nrow(simon)

prop_third_party <- (1 - median(intake$nintendo_first_party_prop, na.rm = T)) * 100
prop_failed_daily_attn <- (sum(daily_raw$bpnsfs_failed_att_check) / nrow(daily_raw)) * 100
prop_failed_biweekly_attn <- (sum(biweekly_raw$bangs_failed_att_check) / nrow(biweekly_raw)) * 100

# no. invalid sessions (shopping, videos, etc.)
prop_invalid_xbox <- (1 - (nrow(xbox_valid_sessions) / nrow(xbox_raw))) * 100
prop_invalid_nintendo <- (1 - (nrow(nintendo_valid_sessions) / nrow(nintendo_raw))) * 100

# no. suspicious sessions (1 - proportion of clean session after removing long hour sessions)
prop_suspicious_xbox <- (1 - (nrow(xbox) / nrow(xbox_valid_sessions))) * 100
prop_suspicious_nintendo <- (1 - (nrow(nintendo) / nrow(nintendo_valid_sessions))) * 100

prop_dropped_timeuse <- (1 - (nrow(time_use) / nrow(time_use_raw))) * 100

n_timeuse_entries_per_day <- time_use |>
    group_by(pid, date) |>
    summarise(n_activites = n()) |>
    summarise(median = median(n_activites, na.rm = TRUE)) |>
    summarise(median = median(median, na.rm = TRUE)) |>
    pull(median)


# This was wrong in the paper? see code chunk below
platforms_count <- str_count(
  replace_na(intake$linked_platforms[intake$qualified], ""),
  ","
) + 1 # single platform won't have a ','
linked_platforms_raw_qual <- intake$linked_platforms[intake$qualified]
platforms_count_fix <- if_else(
  is.na(linked_platforms_raw_qual) | trimws(linked_platforms_raw_qual) == "",
  0,
  str_count(replace_na(linked_platforms_raw_qual, ""), ",") + 1
)
n_one_platform <- sum(platforms_count == 1, na.rm = TRUE)
n_two_platforms <- sum(platforms_count == 2, na.rm = TRUE)
n_three_platforms <- sum(platforms_count >= 3, na.rm = TRUE)

# Steam quality metrics
steam_qc <- read_csv(
  "data/open-play/qc/steam_quality_metrics.csv",
  show_col_types = FALSE
)
prop_steam_hours_exceeding <- steam_qc$prop_hours_exceeding * 100
prop_steam_segments_scaled <- steam_qc$prop_segments_scaled * 100

mean_minutes_reduced <- steam_qc$mean_minutes_reduced
median_minutes_reduced <- steam_qc$median_minutes_reduced
```

There are `{r} n_baseline` qualified participants and `{r} n_participants` took part in the surveys.
Out of the `{r} n_participants`, `{r} n_participants_biweekly` completed the biweekly surveys, and `{r} n_participants_daily` did the daily surveys.

For future reference, the variable `eligible_platforms` contains all platforms that can be used for telemetry.
```{r}
#| label: eligible-platforms
table(intake$eligible_platforms, useNA = 'ifany')
```

The people that did not link their video game services, but were eligible, were initially classified as "Elected not to link".
Additionally, the participants did end up recording their telemetry, as seen in Figure 1 of the Open Play paper.
Therefore, this finding simply indicates the participants at the time elected to not link, but did in the end.

# Intake Data
```{r}
#| label: set-intake-data
intake_participants <- intake |> filter(pid %in% unique(c(daily$pid, biweekly$pid)))
intake_qualified <- intake |> filter(qualified)
```


First, let's get an idea of the participants that took part in the surveys (both daily and biweekly).

```{r}
print(str(intake_participants, list.len = Inf, vec.len = 3))
print(summary(intake_participants))
```
## Missing Data Analysis

```{r}
#| label: intake-na
na_count <- unlist(colSums(is.na(intake_participants)))
na_prop <- unlist(round(colMeans(is.na(intake_participants)) * 100, 2))
intake_na <- data.table(variable = colnames(intake_participants), na_count, na_prop)

tot_na_count <- n_miss(intake_participants)
tot_na_pct <- pct_miss(intake_participants)

intake_na |> arrange(-na_count)
```

## Demographics


```{r}
#| label: na-plot-demographic
# exploring demographic variable missingness
demographic_vars <- c(
  "country",
  "geo_area",
  "age",
  "gender",
  "height",
  "weight",
  "ethnicity",
  "edu_level",
  "employment",
  "marital_status",
  "political_party",
  "dependents",
  "neuro_diagnosed",
  "neuro_identify"
)

intake_na |> filter(variable %in% demographic_vars) |> arrange(-na_count)


aggr(
  intake_participants[, demographic_vars],
  sortVar = TRUE,
  numbers = TRUE,
  prop = FALSE,
  cex.lab = 1,
  cex.axis = 0.5,
  cex.numbers = 0.5
)

vis_miss(intake_participants[, demographic_vars]) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 0, vjust = 0),
    margins = margin(r = 50, b = 5)
  )
```
We can see that the main variables missing data are dependents and neuro_diagnosed.

### Cleaning

I will clean the variables that are not missing, but have lots of varied responses (country, geo_area, age, gender, edu_level, marital_status, political_party, neuro_identify).

#### Country

```{r}
#| label: county-tabs

table(intake_participants$country, useNA = "ifany")
```
There is nothing to be done with country.

#### Geographic Area of Residence

```{r}
#| label: geo_area-tabs

table(intake_participants$geo_area, useNA = "ifany")
```
The US codes look fine, but the UK should be capitalised.
```{r}
#| label: geo_area-clean

intake_participants <- intake_participants |> 
  mutate(
    geo_area_original = geo_area,
    geo_area = str_to_upper(geo_area)
  )
```
That reduced nearly 40 of the possible values.

#### Age

```{r}
#| label: age-tabs

table(intake_participants$age, useNA = "ifany")
```
Nothing to do with age.

#### Gender

```{r}
#| label: gender-tabs

table(intake_participants$gender, useNA = "ifany")
```
I will convert the values to fewer scales for easier analysis (not Man/Woman/Non-binary/Questioning/Other).
The number of transgender and individual non-binary identities are too few to be put as their own response groups, as this may cause bias in future modelling.
Therefore, I will group gender into the following: Male/Female/Non-binary/Other.
Prefer not to say shall be converted to NA, especially as there are only 3.
```{r}
#| label: gender-clean

intake_participants <- intake_participants |> 
  mutate(
    gender_original = gender,
    gender = str_to_lower(gender_original),
    gender = case_when(
      grepl("woman", gender, fixed = TRUE) ~ "Woman",
      grepl("agender", gender, fixed = TRUE) |
        grepl("they|them", gender) |
        grepl("demi", gender, fixed = TRUE) |
        grepl("bi", gender, fixed = TRUE) |
        grepl("gender.*fluid", gender) |
        grepl("non.*binary", gender) ~ "Non-binary",
      grepl("man", gender, fixed = TRUE) ~ "Man",
      gender == "prefer not to say" ~ "Prefer not to say",
      TRUE ~ NA_character_
    )
  )

table(intake_participants$gender, useNA = "ifany")

intake_participants |> select(gender, gender_original) |>
  count(gender, gender_original) |> print()
```

#### Education Level

I will make this data ordinal, as then this could be used for future work.
This will also involve grouping some of the variables together (Highschool diploma/some college)

```{r}
#| label: edu_level-tabs

table(intake_participants$edu_level, useNA = "ifany")
```

Values according to the codebook: Less than a high school diploma or equivalent; High school graduate or equivalent; Some college but no degree; Associate degree; Bachelor's degree; Master's degree, professional degree, or doctoral degree; Prefer not to say

Also note that "prefer not to say" is being encoded as NA.
```{r}
#| label: edu_level-clean

old_edu <- intake_participants$edu_level
intake_participants$edu_level_original <- old_edu

intake_participants <- intake_participants |>
  mutate(
    edu_level = case_when(
      edu_level_original %in% c(
        "Less than a high school diploma or equivalent", 
        "Completed Primary School"
      ) ~ "Below High School",
      edu_level_original %in% c(
        "High school graduate or equivalent",
        "Completed Secondary School"
      ) ~ "High School",
      edu_level_original == "Vocational or Similar" ~ "Vocational or Similar",
      edu_level_original %in% c(
        "Some University but no degree",
        "Some college but no degree"
      ) ~ "Higher education but no degree",
      edu_level_original == "Associate degree" ~ "Associate degree",
      edu_level_original %in% c(
        "Bachelor's degree",
        "University Bachelors Degree",
        "Higher National Certificate, Higher National Diploma, Bachelor's degree, or post-graduate qualifications"
      ) ~ "Bachelors degree",
      edu_level_original %in% c(
        "Master's degree, professional degree, or doctoral degree",
        "Graduate or professional degree (MA, MS, MBA, PhD, etc)"
      ) ~ "Postgraduate",
      edu_level_original == "Prefer not to say" ~ "Prefer not to say",
      TRUE ~ NA_character_
    )
  )

print(intake_participants |> count(edu_level), n = Inf)
```

#### Ethnicity

```{r}
#| label: ethnicity-tabs

table(intake_participants$ethnicity, useNA = "ifany")
```

I will create the following groups: White, Black, Asian, Native Hawaiian, Mixed, Other, American Indian, Prefer not to say (as NA):

```{r}
#| label: ethnicity-clean

ethnicity_old <- intake_participants$ethnicity
intake_participants$ethnicity_original <- ethnicity_old

intake_participants <- intake_participants |> 
  mutate(
    ethnicity = ethnicity_original,
    ethnicity = case_when(
      ethnicity %in% c("White alone", "White") ~ "White",
      ethnicity %in% c(
        "Black",
        "Black, African, Caribbean or Black British",
        "Black or African American alone"
      ) ~ "Black",
      ethnicity %in% c(
        "Asian",
        "Asian alone",
        "Asian or Asian British"
      ) ~ "Asian",
      ethnicity %in% c(
        "Mixed",
        "Mixed or multiple ethnic groups",
        "Two or More Races"
      ) ~ "Multiple",
      ethnicity == "American Indian and Alaska Native alone" ~ "American Indian and Alaska Native",
      ethnicity == "Native Hawaiian and Other Pacific Islander alone" ~ "Native Hawaiian and Other Pacific Islander",
      ethnicity == "Prefer not to say" ~ "Prefer not to say",
      is.na(ethnicity) ~ NA_character_,
      TRUE ~ "Other"
    )
  )

table(intake_participants$ethnicity, useNA = "ifany")

intake_eth_na <- intake_participants |> 
  filter(is.na(ethnicity))

intake_participants |> 
  mutate(ethnicity_na = is.na(ethnicity)) |>
  select(panel_source, ethnicity_na) |>
  table()

eth_na_dates <- intake_participants |> filter(is.na(ethnicity)) |>
  mutate(date = as.Date(date)) |> 
  pull(date)
```

The rest of the ethnicity missingness is when using prolific data, specifically all on the 26th September 2024, except for one which appears to be refusal.
The participants (excluding the one anomaly) were pilot participants and also all part of Prolific.
They were evenly distributed across the countries, but all also linked steam.
This seems like some form of error in responses, indicating MNAR.

These shall be recoded as simply "Not reported".

```{r}
intake_participants <- intake_participants |> 
  mutate(
    ethnicity = if_else(is.na(ethnicity), "Not reported", ethnicity)
  )

table(intake_participants$ethnicity, useNA = "ifany")
```


#### Martial Status
```{r}
#| label: marital-status-tabs

table(intake_participants$marital_status, useNA = "ifany")
```
Everything, except one instance, of "Other" is actually just single, by definition of marital status.

```{r}
#| label: marital-status-clean

old_marital_status <- intake_participants$marital_status

intake_participants$marital_status_original <- old_marital_status

intake_participants <- intake_participants |> 
  mutate(
    marital_status = marital_status_original,
    marital_status = case_when(
      marital_status == "Married" ~ "Married",
      marital_status == "In a domestic partnership or civil union" ~ "In a domestic partnership or civil union",
      grepl("separated", str_to_lower(marital_status), fixed = TRUE) ~ "Separated",
      marital_status == "Divorced" ~ "Divorced",
      marital_status == "Widowed" ~ "Widowed",
      marital_status == "Prefer not to say" ~ "Prefer not to say",
      is.na(marital_status) ~ NA_character_,
      TRUE ~ "Single, never married",
    )
  )

table(intake_participants$marital_status, useNA = "ifany")
```


#### Political Party

```{r}
#| label: political-party-tabs

table(intake_participants$political_party, useNA = "ifany")
```

Future note to self: I could try split the countries into smaller states/regions (wales/scotland).
For now, I will just use the specified parties, not specifically handling the others (socialist, scottish greens).
```{r}
#| label: political-party-clean

old_political_party <- intake_participants$political_party

intake_participants$political_party_original <- intake_participants$political_party

intake_participants <- intake_participants |> 
  mutate(
    political_party = political_party_original,
    political_party = case_when(
      # US
      political_party == "Democratic Party" ~ "Democratic Party",
      political_party == "Republican Party" ~ "Republican Party",
      political_party == "Independent" ~ "Independent",
 
      # UK
      political_party == "Labour Party / Co-operative Party" ~ "Labour Party / Co-operative Party",
      political_party == "Conservative and Unionist Party" ~ "Conservative and Unionist Party",
      political_party == "Liberal Democrats" ~ "Liberal Democrats",
      political_party == "Scottish National Party (SNP)" ~ "Scottish National Party (SNP)",
      
      political_party == "Sinn Féin" ~ "Sinn Féin",
      political_party == "Reform UK" ~ "Reform UK",
      political_party == "Democratic Unionist Party (DUP)" ~ "Democratic Unionist Party (DUP)",
      political_party == "Green Party of England and Wales" ~ "Green Party of England and Wales",
      political_party == "Plaid Cymru" ~ "Plaid Cymru",
      political_party == "No affiliation / Independent" ~ "No affiliation / Independent",
      str_detect(str_to_lower(political_party), paste(
        "none", "unaffiliated", "no political affiliation", "not interested",
        "apolitical", "non political", "i dont care", "non political",
        "i dont pay enough attention to politics to know",
        "not interested in politics",
        sep = "|"
      )) ~ "No affiliation / Independent",
      
      political_party == "Prefer not to say" ~ "Prefer not to say",
      is.na(political_party) ~ NA_character_,

      # "Anarcho communist", "scottish greens", "All bad"
      TRUE ~ "Other"
    )
  )

print(intake_participants |> count(political_party), n = Inf)
```

#### Employment

```{r}
#| label: employment-tabs

table(intake_participants$employment, useNA = "ifany")
```

```{r}
#| label: employment-clean

old_employment <- intake_participants$employment

intake_participants$employment_original <- old_employment

intake_participants <- intake_participants |> 
  mutate(
    employment = employment_original,
    employment = str_to_lower(employment),
    employment = case_when(
      employment == "working full-time" ~ "Working full-time",
      employment == "working part-time" ~ "Working part-time",
      employment == "not currently employed" ~ "Not currently employed",
      employment == "a homemaker or stay-at-home parent" ~ "A homemaker or stay-at-home parent",
      employment == "student" ~ "Student",
      employment == "retired" ~ "Retired",
      employment == "prefer not to say" ~ "Prefer not to say",
      is.na(employment) ~ NA_character_,
      TRUE ~ "Other"
    )
  )

print(intake_participants |> count(employment), n = Inf)
```
The sole NA person self reported 600 minutes (10 hours of games) a week.
I just want to investigate further:
```{r}
na_emp_pid <- intake_participants |> filter(is.na(employment)) |> pull(pid)

na_emp_steam <- steam |> filter(pid == na_emp_pid)

start_survey_date <- min(na_emp_steam$date)
end_period <- start_survey_date + weeks(2)

start_2week_playtime <- na_emp_steam |> 
  filter(date >= start_survey_date & date < end_period)

print(paste("Total playtime between ", start_survey_date, " and ", end_period, ": ", sum(start_2week_playtime$minutes)))

start_2week_playtime |> 
  group_by(date) |> 
  summarise(daily_playtime = sum(minutes)) |> 
  ggplot(aes(x = date, y = daily_playtime)) +
  geom_line() +
  scale_x_date(date_labels = "%a %d", breaks = scales::pretty_breaks(n = 14))
```
Check people's ethnicity from similar locations:
```{r}
na_emp <- intake_participants |> filter(is.na(employment))

intake_participants |> 
  filter(country == na_emp$country, geo_area == na_emp$geo_area)
```
I will therefore treat the response of "NA" to imply they are not currently employed:
```{r}
intake_participants <- intake_participants |> 
  mutate(
    employment = if_else(is.na(employment), "Not currently employed", employment)
  )

table(intake_participants$employment, useNA = "ifany")
```


### Missing Data
The rest of the demographic missingness appears to be individuals not disclosing their information, but I will check these instances for any unusual records to be sure:

```{r}
intake_participants <- intake_participants |> 
  mutate(
    adhd = case_when(
      !is.na(neuro_diag_adhd) ~ "Diagnosed ADHD",
      is.na(neuro_diag_adhd) & !is.na(neuro_iden_adhd) ~ "Self-Identified ADHD",
      TRUE ~ "Neurotypical"
    )
  )
```


#### Dependents

```{r}
vis_miss(intake_participants |> select(demographic_vars, starts_with("care_"))) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 0, vjust = 0),
    margins = margin(r = 50, b = 5)
  )
```
Interestingly, the proportion of missing dependents is inverse to the proportion of people missing care_none (they don't care for anyone).
This makes sense as people who do not care for anyone, will leave dependents blank.
Dependents is based on the care variables:
```{r}
addmargins(xtabs(~ dependents + care_none, data = intake_participants, addNA = TRUE))
```
We can see that everyone who didn't answer dependents, indicated they did not have additional caregiving responsibilities.
This is construct-missingness slightly, and is MAR, as the missingness is heavily dependent on factors such as marital status and age.

I will recode NA dependent participants' "None" if they were answered the care_none variable.
For full transparency in further analysis, I will also add a `imputed` flag to the dataset.

```{r}
intake_participants$dependents_original <- intake_participants$dependents
intake_participants <- intake_participants |> 
  mutate(
    dependents_original = dependents,
    dependents_imputed = if_else(
      is.na(dependents) & !is.na(care_none),
      1, 0
    ),
    dependents = case_when(
      dependents_imputed == 1 ~ "None",
      TRUE ~ dependents
    )
  )
```

```{r}
addmargins(xtabs(~ dependents + care_children, data = intake_participants, addNA = TRUE))
addmargins(xtabs(~ dependents + care_elderly, data = intake_participants, addNA = TRUE))
addmargins(xtabs(~ dependents + care_disability, data = intake_participants, addNA = TRUE))
addmargins(xtabs(~ dependents + care_prefer_not_say, data = intake_participants, addNA = TRUE))
addmargins(xtabs(~ dependents + care_other, data = intake_participants, addNA = TRUE))
```
The final missing 17 rows are all when people responded to "prefer not to say" for care-giving abilities.
I will leave these as is, rather than dropping them as they don't even account for 1% of the data.

#### Height and Weight

Now I will investigate the missing weight and height data:
```{r}
na_weight <- if_else(is.na(intake_participants$weight), 1, 0)
na_height <- if_else(is.na(intake_participants$height), 1, 0)

xtabs(~na_height + na_weight)
```

There is only one record where height is missing and weight isn't, 1840 are not missing, 137 records are missing both.
```{r}
#| label: na weight cross-tabs

check_cols <- demographic_vars[!demographic_vars %in% c("height", "weight", "geo_area")]
for (col in check_cols) {
  cat("Cross-tab for", col, "and na_weight:\n")
  form_string <- paste0("~ ", col, " + na_weight")
  xtable <- xtabs(as.formula(form_string), data = intake_participants, addNA = TRUE)
  na_pct <- round(prop.table(xtable, margin = 1)[, "1"] * 100, 1)
  na_total_pct <- round(xtable[, "1"] / sum(xtable[, "1"]) * 100, 1)
  print(cbind(xtable, pct_na = na_pct, tot_pct_na = na_total_pct))
}
```
Now let's just get a general summary of height and weight:
```{r}
intake_participants |> select(height, weight) |> 
  ggplot(aes(x = weight)) +
  geom_boxplot()

intake_participants |> select(height, weight) |> 
  ggplot(aes(x = height)) +
  geom_boxplot()
```
Let me look at all the outlying values from these columns:
```{r}
# IQR outlier method
height_quartiles <- quantile(intake_participants$height, probs = c(.25, .5, .75), na.rm = TRUE)
weight_quartiles <- quantile(intake_participants$weight, probs = c(.25, .5, .75), na.rm = TRUE)

height_iqr <- IQR(intake_participants$height, na.rm = TRUE)
weight_iqr <- IQR(intake_participants$weight, na.rm = TRUE)

outlying_height_weight <- intake_participants |> 
  filter(
    height < (height_quartiles[1] - 1.5 * height_iqr) |
    height > (height_quartiles[3] + 1.5 * height_iqr) |
    weight < (weight_quartiles[1] - 1.5 * weight_iqr) |
    weight > (weight_quartiles[3] + 1.5 * weight_iqr)
  ) |>
  select(pid, age, country, height, weight, edu_level, marital_status, employment)
```


I can try impute these, or simply leave them in the binary missingness indicator.
The data is missing at an item-level.

```{r}
missing_weight <- intake_participants |> filter(is.na(weight))

missing_weight |> 
  summarise(across(everything(), ~sum(is.na(.x)))) |> 
  pivot_longer(everything(), names_to = "variable", values_to = "na_count") |> 
  arrange(-na_count) |> 
  mutate(pct_na = round(na_count / nrow(missing_weight), 2))
```
It appears that the weight variable has no real relation to any other variables' missingness.
This indicates the data is missing at the item-level.
This seems to be bordering between MAR and MNAR - people can be sensetive to disclosing their weight and height.

Does gender affect weight's missingness:
```{r}
#| label: gender-weight-chisq

# remove the 4 participants with missing gender
clean_data <- subset(intake_participants, !is.na(gender))
clean_data <- clean_data |> filter(gender != "Prefer not to say") # PNTS only has 3 groups
weight_gender_table <- table(clean_data$gender, is.na(clean_data$weight), useNA = "ifany")
chisq.test(weight_gender_table)
weight_gender_table

height_gender_table <- table(clean_data$gender, is.na(clean_data$height), useNA = "ifany")
chisq.test(height_gender_table)
height_gender_table
```
This confirms the weight data is MAR, highlighting nearly 16% of Non-binary participants are missing weight.

I will find variables suitable to be used in multiple imputation (MICE):
```{r}
trialling_cols <- c("pid", "gender", "weight", "height", "age", "ethnicity", "country", "edu_level", "employment", "self_reported_weekly_play")
complete_cases_df <- intake_participants %>%
  filter(!is.na(weight) & !is.na(self_reported_weekly_play)) |>
  mutate(across(c(gender, ethnicity, country, edu_level, employment), as.factor)) |> 
  select(all_of(trialling_cols))

# Fit a temporary diagnostic linear model
vif_model <- lm(weight ~ gender + age + ethnicity + country + edu_level + employment + self_reported_weekly_play, 
                data = complete_cases_df)

# Check VIF scores
vif(vif_model)
```
Features lack multi-collinearity.

```{r}
# pearson correlation for numeric
correlation_data <- intake_participants |> select(age, self_reported_weekly_play, weight) |>
  mutate(na_weight = is.na(weight))
cor(correlation_data[, c("age", "self_reported_weekly_play", "na_weight")], use = "complete.obs", method = "pearson")

# kruskal-wallis test for categoric
for (col in c("country", "gender", "ethnicity", "employment", "edu_level", "marital_status")) {
  # col_weight_formula <- as.formula(paste0("~weight+", col))
  # print(addmargins(xtabs(col_weight_formula, data = intake_participants)))
  print(kruskal.test(as.formula(paste0("weight~", col)), data = intake_participants))
}
```
```{r}
# kruskal-wallis test for categoric
for (col in c("country", "gender", "ethnicity", "employment", "edu_level", "marital_status")) {
  print(ggplot(intake_participants, aes(x = weight)) +
    geom_histogram() +
    facet_wrap(vars(.data[[col]])))
}
```

I will perform multiple imputation with gender, ethnicity, employment, and marital status:

```{r}
#| label: weight-imputation

# check collinearity
impute_subset <- intake_participants |> 
  select(
    pid,
    # targets
    height, weight, 
    # Core biological/demographic predictors
    gender, age, ethnicity, 
    # Geographical/Socioeconomic predictors
    country, edu_level, employment,
  ) |>
  # Convert all textual categorical fields to factors (Required by MICE)
  mutate(
    # remove anomalies quickly - they will be imputed
    weight = if_else(weight < 30, NA_real_, weight),
    height = if_else(height > 400, NA_real_, height),
    
    across(where(is.character), as.factor)
  )

mice_config <- mice(impute_subset, maxit = 0)

# which column was used for imputing the row feature
pred_matrix <- mice_config$predictorMatrix

# force height and weight to be the only features imputed
no_impute_cols <- c("gender", "age", "ethnicity", "country", "edu_level", "employment")
pred_matrix[no_impute_cols, ] <- 0
pred_matrix["pid",] <- 0
pred_matrix[, "pid"] <- 0

# force non-impute cols to have no method
mice_methods <- mice_config$method
mice_methods[] <- ""
mice_methods[c("height", "weight")] <- "pmm"

# 'pmm' (Predictive Mean Matching) handles missing continuous variables seamlessly
imputed_output <- mice(
  data = impute_subset, 
  m = 5, # 5 distinct imputed datasets
  method = mice_methods,
  predictorMatrix = pred_matrix,
  seed = 30
)

map(c("height", "weight"), ~{
  ggmice(imputed_output, aes(x = .imp, y = .data[[.x]])) +
    geom_boxplot() +
    labs(x = "Imputation number")
}) |> 
  wrap_plots()
```
```{r}
print("Means of original data:")
intake_participants |> select(height, weight) |> colMeans(na.rm = TRUE)

print("Means of imputed data:")

imputed_output |>
  mice::complete("all", include = FALSE) |>
  map(select, height, weight) |>
  map(colMeans)
```

Now I will merge these with the dataset:
```{r}
imputed_df <- complete(imputed_output, 1)

imputed_cols <- imputed_df |> 
  select(pid, imputed_height = height, imputed_weight = weight)

intake_participants$height_original <- intake_participants$height
intake_participants$weight_original <- intake_participants$weight

intake_participants <- intake_participants |> 
  left_join(imputed_cols, by = "pid") |> 
  mutate(
    height_imputed = if_else(is.na(height), 1, 0),
    weight_imputed = if_else(is.na(weight), 1, 0),
    height = replace_na(imputed_height),
    weight = replace_na(imputed_weight)
  ) |> 
  select(-imputed_weight, -imputed_height)

summary(intake_participants)
```

# Daily Surveys

```{r} 
daily |>
  mutate(across(where(is.character), as.factor)) |>
  skim() |>
  print()
```
The `stress_events_health` and `stress_events_other` appear to be swapped.
I must check if the number of responses match up to see if the `how_stress` variables are also swapped
```{r} 
addmargins(xtabs(~ stress_events_health + how_stressful_health, data = daily, addNA = TRUE))
addmargins(xtabs(~ stress_events_health + how_stressful_other, data = daily, addNA = TRUE))

addmargins(xtabs(~ stress_events_other + how_stressful_health, data = daily, addNA = TRUE))
addmargins(xtabs(~ stress_events_other + how_stressful_other, data = daily, addNA = TRUE))
```
We can see that when the indicator for health is NA, there are still responses for the level of stress, where there shouldn't be.
This is confirmed by forming the cross-tab with `how_stressful_other`, where there are no responeses when health is NA.
We will therefore swap these:
```{r}
real_stress_health <- daily$stress_events_other
real_how_stress_health <- daily$how_stressful_other

daily_cleaned <- daily |>
  mutate(
    stress_events_other = stress_events_health,
    how_stressful_other = how_stressful_health,
    stress_events_health = real_stress_health,
    how_stressful_health = real_how_stress_health
  )

addmargins(xtabs(~ stress_events_health + how_stressful_health, data = daily_cleaned, addNA = TRUE))
addmargins(xtabs(~ stress_events_health + how_stressful_other, data = daily_cleaned, addNA = TRUE))

addmargins(xtabs(~ stress_events_other + how_stressful_health, data = daily_cleaned, addNA = TRUE))
addmargins(xtabs(~ stress_events_other + how_stressful_other, data = daily_cleaned, addNA = TRUE))
```

# Biweekly Surveys

```{r} 
biweekly |>
  mutate(across(where(is.character), as.factor)) |>
  skim() |>
  print()

skim(biweekly, starts_with("bfi_2"))
```
All of the `bfi` variables are missing.
The `gaming_value` variables are ranged -3 to 3, whereas the codebook says 1-7.
I will recode in accordance with the codebook.
```{r} 
biweekly_cleaned <- biweekly |>
  # -3 to 3 -> 1-7
  mutate(across(starts_with("gaming_value"), function(col) col + 4))

biweekly_cleaned |>
  mutate(across(where(is.character), as.factor)) |>
  skim() |>
  print()
```

There is also a -300 in `recent_session1_duration_min`, which cannot be negative.
```{r}
ggplot(biweekly_cleaned, aes(y = recent_session3_duration_min)) +
  geom_boxplot()

biweekly_cleaned |>
  filter(recent_session3_duration_min >= 180) |> # Q3
  pull(recent_session3_duration_min) |>
  skim()

# only one participant
wrong_session3 <- biweekly_cleaned |>
  filter(recent_session3_duration_min < 0)

biweekly_cleaned |>
  filter(pid == wrong_session3$pid) |>
  pull(recent_session3_duration_min) |>
  skim()
```

Looking at the participant's and the cohort's distribution of this variable's values,
this is within the range (max is 1700, in 93.75 percentile).
I will simply make the value positive
```{r} 
biweekly_cleaned <- biweekly_cleaned |>
  mutate(recent_session3_duration_min = abs(recent_session3_duration_min))
```

# Video Game Telemetry

## Android
```{r}
skim(android)
```
Nothing unusual here.

## iOS
```{r}
skim(ios)
```
Same here

## Xbox
I will work with the `xbox_valid_sessions` as it continues from Ballou's work.
```{r}
skim(xbox |> mutate(across(where(is.character), as.factor)))
```

## Steam
```{r}
skim(steam |> mutate(across(where(is.character), as.factor)))
```

## Metadata
```{r}
skim(metadata |> mutate(across(where(is.character), as.factor)))
```

## Nintendo
```{r}
skim(nintendo |> mutate(across(where(is.character), as.factor)))
```

# Attention (Simon)
```{r}
skim(simon |> mutate(across(where(is.character), as.factor)))

ggplot(simon, aes(x = score_final)) +
  geom_histogram()

negative_simon_score <- simon |>
  filter(score_final < 0)
```
A negative score is allowed as a participant gains one point for each correct response,
and lose a point for every incorrect response.

# Timeuse
```{r}
skim(time_use |> mutate(across(where(is.character), as.factor)))
```
`activity_other` is not missing when the activity recorded in the time use is 'other'.

# Writing the Data
Data to write: survey_intake, survey_daily, survey_biweekly.
```{r} 
write_csv(
  intake_participants,
  file = "data/open-play/clean/survey_intake_participants_final.csv.gz"
)
write_csv(
  daily_cleaned,
  file = "data/open-play/clean/survey_daily_final.csv.gz"
)
write_csv(
  biweekly_cleaned,
  file = "data/open-play/clean/survey_biweekly_final.csv.gz"
)
```
