---
title: "Analysing the final Mapper Graph"
author:
  name: Daniel Agar
format:
  pdf: default
  html:
    toc: true
  docx: default
---

```{r}
#| label: libraries
#| include: false
#| output: false
#| code-summary: "Show code (Load required R packages)"

source("renv/activate.R")
library(tidyverse)
library(MapperAlgo)
library(igraph)
library(jsonlite)
library(knitr) # latex tables
library(skimr)
library(dbscan)
library(janitor)

generate_pdf_figs <- FALSE
theme_set(theme_minimal())
source("examples/open-play/helpers.R")
source("utils.R")

sourceDir("mapper", exclude = "test.R")
source("examples/open-play/mapper_data.R")

mapper_full_data <- mapper_full_data |>
  mutate(adhd = as.factor(adhd))
```

```{r}
#| label: data setup
#| include: false
#| cache: true
#| output: false

# Load the mapper file
grid_dir <- "examples/open-play/grid_searches"
mapper_file <- paste(
  # "shortlisted",
  # "more_telem_pca12_zscale",
  # "more_telem_pca12_final_indfix_zscale",
  "more_telem_pca12_final_zscale",
  # "more_telem_pca12_final_covextension_int10_ov40_dbscan_eps3_minPts3.json",
  "more_telem_pca12_final_covextension_int10_ov40_dbscan_eps3_minPts5.json",
  # "more_telem_pca12_final_indfix_covextension_int10_ov40_dbscan_eps0.4_minPts5.json",
  # "more_telem_pca12_covstride_int5_ov50_kmeans_max_kmeans_clusters2.json",
  # "more_telem_pca12_covstride_int10_ov50_dbscan_eps0.05_minPts3.json",
  # "more_telem_pca12_covstride_int5_ov50_dbscan_eps0.4_minPts34.json",
  sep = "/"
)
mapper_graph_data <- fromJSON(here::here(paste(grid_dir, mapper_file, sep = "/")))

any_adhd <- mapper_full_data$any_adhd
```

```{r} 
#| label: data scaling
#| output: false
scaler <- "zscale"

# Min-Max Normalisation (commonly used in Kepler Mapper)
min_max_norm <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

if (scaler == "minmax") {
  mapper_scaled <- mapper_data |>
    mutate(across(everything(), min_max_norm))
} else if (scaler == "zscale") {
  mapper_scaled <- mapper_data |>
    mutate(across(everything(), ~ scale(.)[, 1]))
} else {
  stop("No valid scaler selected")
}

mapper_pca <- prcomp(mapper_data, center = TRUE, scale = TRUE)
# print(xtable::xtable(round(cor(cbind(
#   mapper_scaled,
#   pc1 = mapper_pca$x[, 1],
#   pc2 = mapper_pca$x[, 2]
# ))[, c("pc1", "pc2")], 2),
# floating = FALSE, latex.environments = NULL, booktabs = TRUE))
```

```{r}
#| label: dimension reduction trials
#| include: false
#| eval: false

library(uwot)
library(Rtsne)

set.seed(30)
mapper_umap <- umap2(
  mapper_data,
  n_components = 2,
  scale = TRUE,
  n_neighbors = 10,
  min_dist = 0.001
)
mapper_pca <- prcomp(mapper_data, center = TRUE, scale = TRUE)
set.seed(30)
mapper_tsne <- Rtsne(
  mapper_data,
  dims = 2,
  pca_center = TRUE,
  pca_scale = TRUE
)
mapper_dists <- dist(mapper_scaled)
mapper_mds <- cmdscale(mapper_dists, k = 2)

adhd_group <- mapper_full_data$adhd

dim_red_data <- bind_rows(
  tibble(
    method = "pca",
    dim1 = mapper_pca$x[, 1],
    dim2 = mapper_pca$x[, 2],
    adhd = adhd_group
  ),
  tibble(
    method = "tsne",
    dim1 = mapper_tsne$Y[, 1],
    dim2 = mapper_tsne$Y[, 2],
    adhd = adhd_group
  ),
  tibble(
    method = "umap",
    dim1 = mapper_umap[, 1],
    dim2 = mapper_umap[, 2],
    adhd = adhd_group
  ),
  tibble(
    method = "mds",
    dim1 = mapper_mds[, 1],
    dim2 = mapper_mds[, 2],
    adhd = adhd_group
  )
)

if (generate_pdf_figs) pdf("mapper_analysis_figs/dimension_reductions_filters.pdf")
ggplot(dim_red_data, aes(x = dim1, y = dim2, colour = adhd)) +
  geom_point(size = 0.3, alpha = 0.6, shape = 1) +
  facet_wrap(vars(method), scale = "free")
if (generate_pdf_figs) dev.off()
```

# Mapper Graph
```{r}
#| label: create mapper

# 10 intervals, 50% overlap
# DBSCAN(0.4, 26)
mapper <- MapperAlgo(
  original_data = mapper_scaled,
  filter_values = mapper_pca$x[, 1:2],
  percent_overlap = 40,
  methods = "dbscan",
  method_params = list(eps = 3, minPts = 5),
  cover_type = "stride",
  intervals = 10,
  num_cores = 10
)
# apply index fix
# mapper$points_in_vertex <- lapply(mapper$points_in_vertex, function(x) x - 1)
```

```{r} 
#| label: dbscan params figure

# histogram of the euclidean distances
mapper_dists <- dist(mapper_scaled)
mapper_dists <- as.numeric(mapper_dists)

if (generate_pdf_figs) pdf("mapper_analysis_figs/mapper_scaled_dists_hist.pdf")
ggplot(data.frame(distance = mapper_dists), aes(x = distance)) +
  geom_histogram(color = "black", bins = 20) +
  geom_vline(xintercept = 3, colour = "red") +
  scale_x_continuous(breaks = seq(0, 25, 5))
if (generate_pdf_figs) dev.off()

# List of clustering methods and their respective hyperparameter sets
# k ~ 2 x dimensions
if (generate_pdf_figs) pdf("mapper_analysis_figs/knn_distplot_zscale_k5_scaled_data.pdf")
kNNdistplot(x = mapper_scaled, k = 5) # minPts~3 is good
abline(h = 3, col = "red")
if (generate_pdf_figs) dev.off()
```
Looking at these plots and changing k in the 'kNNdistplot' gives eps~3 a good choice

```{r}
#| label: graph any adhd

# important keep the seed
set.seed(2)
if (generate_pdf_figs) pdf("mapper_analysis_figs/final_mapper_any_adhd.pdf")
create_mapper_graph(
  mapper_obj = mapper_graph_data,
  colourisation = mapper_full_data$any_adhd,
  original_data = mapper_data,
  graph_layout = layout_with_fr,
  legend = TRUE,
  legend_title = "Proportion of ADHD Participants",
  legend_tick_num = 5
)
if (generate_pdf_figs) dev.off()
```

```{r}
#| label: graph colourisation

# bangs comp sat: strongest correlation with PC1
set.seed(2)
if (generate_pdf_figs) pdf("mapper_analysis_figs/final_mapper_bangs_comp_sat.pdf")
create_mapper_graph(
  mapper_obj = mapper_graph_data,
  colourisation = mapper_data$bangs_comp_sat,
  original_data = mapper_data,
  graph_layout = layout_with_fr,
  legend = TRUE,
  legend_title = "BANGS Competence Satisfaction",
  legend_tick_num = 5
)
if (generate_pdf_figs) dev.off()

set.seed(2)
# total wave minutes: strongest correlation with PC2
if (generate_pdf_figs) pdf("mapper_analysis_figs/final_mapper_total_wave_minutes.pdf")
create_mapper_graph(
  mapper_obj = mapper_graph_data,
  colourisation = mapper_data$total_wave_minutes,
  original_data = mapper_data,
  graph_layout = layout_with_fr,
  legend = TRUE,
  legend_title = "Total wave playtime (minutes)",
  legend_tick_num = 5
)
if (generate_pdf_figs) dev.off()
```


```{r}
#| label: game metadata and isolated components

isolated_components <- find_isolated_components(mapper_graph_data, mapper_full_data)

# join node and component to the data
graph_full_data <- mapper_full_data |>
  add_column(node = 0, component = 0)

for (row in rownames(isolated_components)) {
  row_data <- isolated_components[row, ]
  node_points <- mapper_graph_data$points_in_vertex[[row_data$node]]
  graph_full_data[node_points, "node"] <- row_data$node
  graph_full_data[node_points, "component"] <- row_data$component
}


interested_cols <- c(
  "pid", "wave", "adhd",
  "wemwbs_total", "life_sat", "aff_val",
  "bangs_auto_sat", "bangs_auto_frus", "bangs_comp_sat",
   "bangs_comp_frus", "bangs_rel_sat", "bangs_rel_frus",
  "gaming_value_sum", "total_wave_minutes", "max_binge_minutes",
  "binge_day_of_week", "mean_mobile", "mean_nintendo",
  "mean_steam", "mean_xbox", "score_mean", "atten_score_mean"
)

print("Full data:")
print(skim(mapper_full_data))

# create full game data table
all_game_telemetry <- bind_rows(
  steam |>
    mutate(platform = "Steam") |>
    select(
      pid,
      platform,
      session_start = approximate_session_start,
      session_end = approximate_session_end,
      duration = minutes,
      title_id
    ),
  nintendo |>
    mutate(platform = "Nintendo") |>
    select(
      pid,
      platform,
      session_start,
      session_end,
      duration,
      title_id
    ),
  # mobile combined
  bind_rows(ios, android) |>
    mutate(platform = "Mobile", title_id = NA_character_) |>
    mutate(duration = sum(minutes), .by = c(pid, day_local)) |>
    select(
      pid,
      platform,
      session_start = day_local,
      session_end = day_local,
      duration,
      title_id
    )
) |>
  left_join(metadata, by = join_by(platform, title_id == original_name))

xbox_game_telemetry_metadata <- xbox |>
  mutate(platform = "Xbox") |>
  select(
    pid,
    platform,
    session_start,
    session_end,
    duration,
    xbox_title_id = title_id
  ) |>
  left_join(metadata, by = join_by(
    platform,
    xbox_title_id == original_name
  ))

# TODO: manual game renaming
game_sessions <- bind_rows(
  all_game_telemetry,
  xbox_game_telemetry_metadata
) |>
  mutate(
    primary_genre = case_when(
      platform == "Xbox" ~ xbox_primary_genre,
      platform == "Mobile" ~ "Mobile",
      is.na(genres) ~ NA_character_,
      TRUE ~ str_split_i(genres, ",", 1)
    ),
    secondary_genre = case_when(
      platform == "Xbox" ~ xbox_subgenre,
      platform == "Mobile" ~ "Mobile",
      is.na(genres) ~ NA_character_,
      TRUE ~ trimws(str_split_i(genres, ",", 2))
    )
  ) |>
  select(
    pid, title_id, platform,
    session_start, session_end, duration,
    starts_with("igdb_"), match_confidence,
    platforms, primary_genre, secondary_genre
  )

mapper_data_participants <- biweekly_mapper_data |>
  left_join(intake_participants, by = "pid")
```

```{r}
#| label: game genre colour scales

game_sessions |> count(primary_genre, sort = TRUE) |> print(n = Inf)
game_sessions |> count(secondary_genre, sort = TRUE) |> print(n = Inf)

# proportions of a mapper row's telemetry playing different genres
genre_maps <- tribble(
  ~from, ~to,
  c("Card & Board Game", "Card + Board"), "Card & Board",
  c("Platform", "Platformer"), "Platformer",
  c("Puzzle", "Puzzle + Trivia", "Quiz/Trivia"), "Puzzle & Trivia",
  c("Racing", "Racing + Flying"), "Racing",
  c("Role Playing", "Role-playing (RPG)"), "Role-Playing (RPG)",
  c("Simulation", "Simulator"), "Simulation",
  c("Sport", "Sports"), "Sports",
  c("Real Time Strategy (RTS)", "Strategy", "Tactical"), "Strategy",
  c("Multi-Player Online Battle Arena"), "MOBA"
)

biweekly_genres <- mapper_full_data |>
  select(pid, wave, linked_platforms) |>
  left_join(wave_daily_telemetry, by = join_by(pid, wave)) |>
  select(
    pid, wave, linked_platforms,
    wave_start_date = min_twoweek_date,
    wave_end_date = survey_date
  ) |>
  distinct() |>
  left_join(game_sessions, by = join_by(
    pid,
    wave_start_date <= session_start,
    wave_end_date > session_end
  )) |>
  mutate(
    clean_genre = replace_values(
      primary_genre,
      from = genre_maps$from,
      to = genre_maps$to
    )
  ) |>
  summarise(
    wave_minutes_played = sum(duration, na.rm = TRUE),
    .by = c(pid, wave, clean_genre)
  ) |>
  complete(
    nesting(pid, wave),
    clean_genre = unique(na.omit(clean_genre)),
    fill = list(wave_minutes_played = 0)
  ) |>
  filter(!is.na(clean_genre)) |>
  mutate(
    total_wave_minutes = sum(wave_minutes_played),
    genre_wave_prop = if_else(
      total_wave_minutes > 0,
      wave_minutes_played / total_wave_minutes,
      0
    ),
    .by = c(pid, wave)
  ) |>
  select(-total_wave_minutes)

biweekly_genre_wider <- biweekly_genres |>
  select(pid, wave, clean_genre, genre_wave_prop) |>
  pivot_wider(
    names_from = clean_genre,
    names_prefix = "prop",
    values_from = genre_wave_prop,
    values_fill = 0
  ) |>
  clean_names()

mapper_full_data <- mapper_full_data |>
  left_join(biweekly_genre_wider)
```

## Graph Data

```{r}
node_sizes <- lapply(mapper_graph_data$points_in_vertex, length)
unlist_node_sizes <- unlist(node_sizes)
node_sizes[order(unlist(node_sizes))]
```

## Clustered Data
```{r}
#| label: mapper data points

# checking data for index issues
raw_indices <- unlist(mapper_graph_data$points_in_vertex)
is_zero_indexed <- min(raw_indices, na.rm = TRUE) == 0
assigned_rows <- unique(if (is_zero_indexed) raw_indices + 1 else raw_indices)
total_rows    <- nrow(mapper_full_data)
unassigned_rows <- setdiff(seq_len(total_rows), assigned_rows)
message(sprintf(
  "Total Rows: %d | Assigned: %d | Excluded (DBSCAN Noise / Outliers): %d",
  total_rows, length(assigned_rows), length(unassigned_rows)
))
noise_participants <- mapper_full_data[unassigned_rows, ]
clustered_participants <- mapper_full_data[-unassigned_rows,]
mapper_full_data$noise <- FALSE
mapper_full_data[unassigned_rows, "noise"] <- TRUE
```
The nature of DBSCAN excludes noise.
The algorithm classed 166 of the 5836 points as noise.

```{r}
#| label: dbscan noise pca plot

if (generate_pdf_figs) {
  pdf("mapper_analysis_figs/noise_pca_data.pdf")
  par(
    mar = c(3.2, 3.2, 0.8, 0.8), # bottom, left, top, right
    mgp = c(1.8, 0.5, 0), # distance for c(axis title, tick labels, tick line)
    tcl = -0.3 # slightly shorter tick marks
  )
}
plot(
  mapper_pca$x[, 1:2],
  col = "black",
  pch = 16,
  cex = 0.6,
  main = NULL,
  xlab = "PC1",
  ylab = "PC2"
)
points(
  mapper_pca$x[unassigned_rows, 1:2, drop = FALSE],
  col = "red",
  pch = 19,
  cex = 1.2
)
if (generate_pdf_figs) dev.off()
```

The noise appears to be spread more over PC2, which correlates most with telemetry data.
I will check the distributions of the telemetry data.

```{r}
#| label: plotting telemetry distributions

telemetry_cols <- c(
  "total_wave_minutes", "total_steam_minutes",
  "total_xbox_minutes", "total_mobile_minutes",
  "total_nintendo_minutes", "max_binge_minutes",
  "days_played"
)
mapper_full_data |>
  pivot_longer(
    cols = all_of(telemetry_cols),
    names_to = "metric",
    values_to = "value"
  ) |>
  ggplot(aes(x = value, fill = as.factor(noise))) +
  geom_density(alpha = 0.3) +
  facet_wrap(vars(metric), scales = "free") +
  labs(
    x = "Value",
    y = "Density",
    fill = "Noise",
  )

wellbeing_cols <- c(
  "aff_val", "wemwbs_total",
  "bangs_auto_sat", "bangs_auto_frus",
  "bangs_comp_sat", "bangs_comp_frus",
  "bangs_rel_sat", "bangs_rel_frus",
  "gaming_value_sum"
)
mapper_full_data |>
  pivot_longer(
    cols = all_of(wellbeing_cols),
    names_to = "metric",
    values_to = "value"
  ) |>
  ggplot(aes(x = value, fill = as.factor(noise))) +
  geom_density(alpha = 0.3) +
  facet_wrap(vars(metric), scales = "free") +
  labs(
    x = "Value",
    y = "Density",
    fill = "Noise"
  )
```

The outliers appear to show less general playtime and less frustrations with games.
However, we must consider the significant difference in sample sizes.

```{r}
mat_scaled <- as.matrix(mapper_scaled)
feature_blame <- purrr::map_dfr(unassigned_rows, function(i) {
  # squared distances
  diffs_sq <- sweep(mat_scaled[-i, , drop = FALSE], 2, mat_scaled[i, ], "-")^2
  total_dists <- sqrt(rowSums(diffs_sq))
  
  nearest_idx <- which.min(total_dists)
  sq_contrib <- diffs_sq[nearest_idx, ]
  
  tibble(
    noise_point_id = i,
    closest_dist = total_dists[nearest_idx],
    top_driver = names(which.max(sq_contrib)),
    top_driver_share = max(sq_contrib) / sum(sq_contrib)
  )
})

# See which features most frequently push points outside the eps radius
table(feature_blame$top_driver)

# assessing affective valence differences
# testing variances (homoscedasticity)
var.test(clustered_participants$aff_val, noise_participants$aff_val)
aff_val_noise_test <- t.test(clustered_participants$aff_val, noise_participants$aff_val, var.equal = FALSE)
print(aff_val_noise_test)
```

## Isolated Components
The graph shows one large structure,
which shall be the topic of the analysis.
There is however one isolated point in the space.
<!-- Component 2: 1 node, 1 row -->
```{r}
#| label: component 2

comp2_nodes <- isolated_components |>
  filter(component == 2) |>
  pull(node)
unique_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(comp2_nodes)]))
comp2_data <- mapper_full_data[unique_rows, ]
glimpse(comp2_data)

if (generate_pdf_figs) {
  pdf("mapper_analysis_figs/comp2_pca_data.pdf")
  par(
    mar = c(3.2, 3.2, 0.8, 0.8), # bottom, left, top, right
    mgp = c(1.8, 0.5, 0), # distance for c(axis title, tick labels, tick line)
    tcl = -0.3 # slightly shorter tick marks
  )
}
plot(
  mapper_pca$x[, 1:2],
  col = "black",
  pch = 16,
  cex = 0.6,
  main = NULL,
  xlab = "PC1",
  ylab = "PC2"
)
points(
  mapper_pca$x[unique_rows, 1:2, drop = FALSE],
  col = "red",
  pch = 19,
  cex = 1.2
)
if (generate_pdf_figs) dev.off()

# mapper graph visualised by age
create_mapper_graph(
  mapper_obj = mapper_graph_data,
  colourisation = mapper_data_participants$age,
  original_data = mapper_data,
  graph_layout = layout_with_fr,
  legend = TRUE,
  legend_title = "Participant Age",
  legend_tick_num = 5
)
```

1 row of data, Neurotypical, 0 affective valence, 7 wemwbs, sat=3, frus=21 for all.
Gaming sum avg=20, no impact is calculated at 20 (4 times 5).
Played for average of 5 hours each day they played, largest binge is 810 minutes.
This is only one out of their 4 waves they submitted (pid=p3011712557).

The games they played need investigating.
```{r} 
#| label: checking comp 2 games

comp2_pid_info <- comp2_data |>
  select(pid, wave, linked_platforms) |>
  left_join(wave_daily_telemetry, by = join_by(pid, wave)) |>
  select(
    pid, wave, linked_platforms,
    wave_start_date = min_twoweek_date,
    wave_end_date = survey_date
  ) |>
  distinct()

# get dates valid for the wave
comp2_games <- comp2_pid_info |>
  # steam
  left_join(game_sessions, by = join_by(
    pid,
    wave_start_date <= session_start,
    wave_end_date > session_end
  )) |>
  mutate(
    minutes_played = sum(duration),
    .by = c(pid, title_id)
  ) |>
  select(-duration) |>
  distinct(pid, title_id, .keep_all = TRUE)

comp2_games
```
Game data reveals the participant's data associated with this node played online multiplayer shooter games (world of tanks and counter strike) for majroity of the time.
This explains the high frustrations as these are competitive games and can be toxic environments.


## Main Structure
```{r}
#| label: main structure description

main_nodes <- isolated_components |>
  filter(component == 1) |>
  pull(node)

unique_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(main_nodes)]))
main_data <- mapper_full_data[unique_rows, ]

row_node_lookup <- tibble(
  node = seq_along(mapper_graph_data$points_in_vertex),
  row_id = mapper_graph_data$points_in_vertex
) |>
  unnest_longer(row_id) |>
  summarise(nodes = list(unique(node)), .by = "row_id")

main_data <- main_data |>
  mutate(row_id = row_number()) |>
  left_join(row_node_lookup, by = "row_id") |>
  select(-row_id)

skim(main_data)
```

The main structure consists of 90 nodes and 5669 rows of data.
There are small tails on the end corners.
There appears to be an increase in ADHD proportions at the corners.
As demonstrated through the PCA correlations, we see positive experiences and wellbeing decrease as the the node size decreases.

We will first examine the longest tail,
which hints at being formed from low BANGS competence satisfaction scale and other low wellbeing scores.
This hinted at from the mapper graph from before.

### Long Tail

This also includes the base of the tail, for better understanding.
We can guess the tail is related to low scores on wellbeing and BANGS.
```{r}
#| label: long tail

tail_nodes <- c(49, 37, 27, 17)
unique_tail_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(tail_nodes)]))
tail_data <- mapper_full_data[unique_tail_rows, ] |> 
  mutate(pid = as.factor(pid))
skim(tail_data)
```

This tail contains 33 rows of data.
8 data points are from Diagnosed ADHD participants.
The mean affective valence of the flare is 24, but has high variability (sd=20.4)
The wemwbs scores are lower, but has one higher (27).
50% of the tail has BANGS relatedness satisfaction less than or below 6.
They also have low BANGS competence satisfaction.
They experience very high frustration (remember: the average for a construct is 9) of mean values between 13 and 15.

```{r}
#| label: checking tail games

tail_pid_info <- tail_data |>
  select(pid, wave, linked_platforms) |>
  left_join(wave_daily_telemetry, by = join_by(pid, wave)) |>
  select(
    pid, wave, linked_platforms,
    wave_start_date = min_twoweek_date,
    wave_end_date = survey_date
  ) |>
  distinct()

# get dates valid for the wave
tail_games <- tail_pid_info |>
  # steam
  left_join(game_sessions, by = join_by(
    pid,
    wave_start_date <= session_start,
    wave_end_date > session_end
  )) |>
  mutate(
    minutes_played = sum(duration),
    .by = c(pid, title_id)
  ) |>
  select(-duration) |>
  distinct(pid, title_id, .keep_all = TRUE)

tail_games |> summarise(total_played = sum(minutes_played), .by = "primary_genre")
```

## Higher ADHD Proportions
We highlight nodes 91, 85, and 78 as nodes among the general population with higher ADHD proportions.

```{r}
#| label: node 91

unique_node91_rows <- unique(unlist(mapper_graph_data$points_in_vertex[91]))
node91_data <- mapper_full_data[unique_node91_rows, ] |> 
  mutate(pid = as.factor(pid))
skim(node91_data)
```

```{r}
#| label: node 85

unique_node85_rows <- unique(unlist(mapper_graph_data$points_in_vertex[85]))
node85_data <- mapper_full_data[unique_node85_rows, ] |> 
  mutate(pid = as.factor(pid))
skim(node85_data)
```

```{r}
#| label: node 78

unique_node78_rows <- unique(unlist(mapper_graph_data$points_in_vertex[78]))
node78_data <- mapper_full_data[unique_node78_rows, ] |> 
  mutate(pid = as.factor(pid))
skim(node78_data)
```

```{r}
#| label: node 71

unique_node71_rows <- unique(unlist(mapper_graph_data$points_in_vertex[71]))
node71_data <- mapper_full_data[unique_node71_rows, ] |> 
  mutate(pid = as.factor(pid), adhd = as.factor(adhd))
skim(node71_data)
```

## ADHD bottom edge
This section explores the bottom edge near the flare of ADHD proportions.
This contains a small portion of data (min node size 1, max node size 13).
We initially see:
slightly higher attention scores,
higher playtime and binge minutes (expected based on the nature of the filter functions),

```{r}
small_flare_nodes <- c(8, 18, 9, 19)
unique_small_flare_rows <- unique(unlist(mapper_graph_data$points_in_vertex[small_flare_nodes]))
small_flare_data <- mapper_full_data[unique_small_flare_rows, ] |> 
  mutate(pid = as.factor(pid), adhd = as.factor(adhd))
skim(small_flare_data)
```

How many participants are there in this area with ADHD?

## Mobile telemetry
Nodes 22 and 23 have the most.

```{r}
mobile_nodes <- c(22, 23)
unique_mobile_rows <- unique(unlist(mapper_graph_data$points_in_vertex[mobile_nodes]))
mobile_data <- mapper_full_data[unique_mobile_rows, ] |> 
  mutate(pid = as.factor(pid), adhd = as.factor(adhd))
skim(mobile_data)
```

## Small Flares
We do see some little flares beginning:

```{r}
small_flares_nodes <- c(25, 26, 15, 16, 14, 5, 6)
small_flares_rows <- unique(unlist(mapper_graph_data$points_in_vertex[small_flares_nodes]))
small_flares_data <- mapper_full_data[small_flares_rows, ] |>
  mutate(pid = as.factor(pid), adhd = as.factor(adhd))

small_flares_pid_info <- small_flares_data |>
  select(pid, wave, linked_platforms) |>
  left_join(wave_daily_telemetry, by = join_by(pid, wave)) |>
  select(
    pid, wave, linked_platforms,
    wave_start_date = min_twoweek_date,
    wave_end_date = survey_date
  ) |>
  distinct()

# get dates valid for the wave
small_flares_games <- small_flares_pid_info |>
  # steam
  left_join(game_sessions, by = join_by(
    pid,
    wave_start_date <= session_start,
    wave_end_date > session_end
  )) |>
  mutate(
    minutes_played = sum(duration),
    .by = c(pid, title_id)
  ) |>
  select(-duration) |>
  distinct(pid, title_id, .keep_all = TRUE)

small_flares_games |> summarise(total_played = sum(minutes_played), .by = "primary_genre")
skim(small_flares_data)
```

Small flare of nodes 8, 9, 18, 19:

```{r}
second_flare_nodes <- c(8, 9, 18, 19)
second_flare_rows <- unique(unlist(mapper_graph_data$points_in_vertex[second_flare_nodes]))
second_flare_data <- mapper_full_data[second_flare_rows, ] |>
  mutate(pid = as.factor(pid), adhd = as.factor(adhd))

second_flare_pid_info <- second_flare_data |>
  select(pid, wave, linked_platforms) |>
  left_join(wave_daily_telemetry, by = join_by(pid, wave)) |>
  select(
    pid, wave, linked_platforms,
    wave_start_date = min_twoweek_date,
    wave_end_date = survey_date
  ) |>
  distinct()

# get dates valid for the wave
second_flare_games <- second_flare_pid_info |>
  # steam
  left_join(game_sessions, by = join_by(
    pid,
    wave_start_date <= session_start,
    wave_end_date > session_end
  )) |>
  mutate(
    minutes_played = sum(duration),
    .by = c(pid, title_id)
  ) |>
  select(-duration) |>
  distinct(pid, title_id, .keep_all = TRUE)

second_flare_games |> summarise(total_played = sum(minutes_played), .by = "primary_genre")
skim(second_flare_data)
```

# Visualisations

```{r}
#| label: generate any chosen graph

configs <- tribble(
  ~filename, ~data, ~legend_title,
  "final_mapper_aff_val.pdf", mapper_data$aff_val, "Affective Valence",
  "final_mapper_atten_score.pdf", mapper_data$atten_score_mean, "Simon attention task average",
  "final_mapper_bangs_auto_sat.pdf", mapper_data$bangs_auto_sat, "BANGS Autonomy Satisfaction",
  "final_mapper_bangs_comp_sat.pdf", mapper_data$bangs_comp_sat, "BANGS Competence Satisfaction",
  "final_mapper_bangs_rel_frus.pdf", mapper_data$bangs_rel_frus, "BANGS Relatedness Frustration",
  "final_mapper_bangs_rel_sat.pdf", mapper_data$bangs_rel_sat, "BANGS Relatedness Satisfaction",
  "final_mapper_days_played.pdf", mapper_data$days_played, "No. days played videogames",
  "final_mapper_gaming_value_sum.pdf", mapper_data$gaming_value_sum, "Videogame's real-world influence",
  "final_mapper_max_session.pdf", mapper_data$max_binge_minutes, "Longest gaming session",
  "final_mapper_mobile_telemetry.pdf", mapper_data$total_mobile_minutes, "Total mobile playtime (minutes)",
  "final_mapper_nintendo_telemtry.pdf", mapper_data$total_nintendo_minutes, "Total Nintendo playtime (minutes)",
  "final_mapper_xbox_telemetry.pdf", mapper_data$total_xbox_minutes, "Total Xbox playtime (minutes)",
  "final_mapper_prop_action_adventure.pdf", mapper_full_data$prop_action_adventure, "Action & Adventure playtime proportion",
  "final_mapper_prop_adventure.pdf", mapper_full_data$prop_adventure, "Adventure playtime proportion",
  "final_mapper_prop_rpg.pdf", mapper_full_data$prop_role_playing_rpg, "Role-Playing (RPG) playtime proportion",
  "final_mapper_prop_shooter.pdf", mapper_full_data$prop_shooter, "Shooter playtime proportion",
  "final_mapper_prop_simulation.pdf", mapper_full_data$prop_simulation, "Simulation playtime proportion",
  "final_mapper_prop_sports.pdf", mapper_full_data$prop_simulation, "Sports playtime proportion",
  "final_mapper_total_wemwbs_total.pdf", mapper_data$wemwbs_total, "Total WEMWBS measure"
)

for (row in seq_len(nrow(configs))) {
  filename <- configs$filename[row]
  colour_data <- configs$data[[row]]
  title <- configs$legend_title[row]
  
  if (generate_pdf_figs) pdf(paste("mapper_analysis_figs", filename, sep = "/"))
  set.seed(2)
  create_mapper_graph(
    mapper_obj = mapper_graph_data,
    colourisation = colour_data,
    original_data = mapper_data,
    graph_layout = layout_with_fr,
    legend = TRUE,
    legend_title = title,
    legend_tick_num = 5
  )
  if (generate_pdf_figs) dev.off()
}
```

# Descriptive Statistics

```{r}
#| label: descriptive statistics

library(scales)

# telemetry distributions
telemetry_cols <- c(
  "total_wave_minutes", "total_steam_minutes",
  "total_xbox_minutes", "total_mobile_minutes",
  "total_nintendo_minutes", "max_binge_minutes",
  "days_played"
)

telemetry_labels <- c(
  total_mobile_minutes = "Mobile Playtime (min)",
  total_nintendo_minutes = "Nintendo Playtime (min)",
  total_steam_minutes = "Steam Playtime (min)",
  total_xbox_minutes = "Xbox Playtime (min)",
  total_wave_minutes = "Total Wave Playtime (min)",
  days_played = "Days Active",
  max_binge_minutes = "Longest Session (min)"
)

telemetry_long <- mapper_full_data |>
  pivot_longer(
    cols = all_of(telemetry_cols),
    names_to = "metric",
    values_to = "value"
  )

ggplot(telemetry_long, aes(x = value, fill = adhd, colour = adhd)) +
  geom_density(alpha = 0.25, linewidth = 0.45) +
  facet_wrap(
    vars(metric),
    scales = "free",
    labeller = as_labeller(telemetry_labels, default = label_value)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = NULL,
    y = "Density",
    fill = "ADHD Status",
    colour = "ADHD Status"
  ) +
  theme(
    # bottom legend
    legend.position = "bottom",
    legend.box.margin = margin(t = -4, b = 0),
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(face = "bold", size = 7.5),
    legend.text = element_text(size = 7),
    legend.key.size = unit(0.35, "cm"),
    legend.background = element_blank()
  )
```

```{r}
#| label: statistically testing adhd max_binge_minutes

library(rstatix)
library(dunn.test)

desc_summary <- mapper_full_data |>
  group_by(adhd) |>
  summarise(
    N = n(),
    Median = median(max_binge_minutes, na.rm = TRUE),
    IQR = IQR(max_binge_minutes, na.rm = TRUE),
    Mean = mean(max_binge_minutes, na.rm = TRUE),
    SD = sd(max_binge_minutes, na.rm = TRUE),
    .groups = "drop"
  )
print(desc_summary)

# non-parametric test
kw_test <- kruskal.test(max_binge_minutes ~ adhd, data = mapper_full_data)
print(kw_test)

# effect Size (eta-squared for Kruskal-Wallis)
kw_eff <- mapper_full_data |> 
  kruskal_effsize(max_binge_minutes ~ adhd)
print(kw_eff)

# post-hoc pairwise comparisons (Dunn's Test with Holm-Bonferroni Correction)
posthoc_dunn <- dunn.test(
  x = mapper_full_data$max_binge_minutes,
  g = mapper_full_data$adhd,
  method = "holm",
  kw = FALSE
)
```

```{r}
#| label: wellbeing plot

wellbeing_cols <- c(
  "aff_val", "wemwbs_total",
  "bangs_auto_sat", "bangs_auto_frus",
  "bangs_comp_sat", "bangs_comp_frus",
  "bangs_rel_sat", "bangs_rel_frus",
  "gaming_value_sum"
)
wellbeing_labels <- c(
  aff_val = "Affective Valence",
  wemwbs_total = "Total WEMWBS",
  bangs_auto_sat = "BANGS Autonomy Satisfaction",
  bangs_auto_frus = "BANGS Autonomy Frustration",
  bangs_comp_sat = "BANGS Competence Satisfaction",
  bangs_comp_frus = "BANGS Competence Frustration",
  bangs_rel_sat = "BANGS Relatedness Satisfaction",
  bangs_rel_frus = "BANGS Relatedness Frustration",
  gaming_value_sum = "Gaming real-world influence"
)
wellbeing_long <- mapper_full_data |>
  pivot_longer(
    cols = all_of(wellbeing_cols),
    names_to = "metric",
    values_to = "value"
  )

ggplot(wellbeing_long, aes(x = value, fill = adhd, colour = adhd)) +
  geom_density(alpha = 0.25, linewidth = 0.45) +
  facet_wrap(
    vars(metric),
    scales = "free",
    labeller = as_labeller(wellbeing_labels, default = label_value)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = NULL,
    y = "Density",
    fill = "ADHD Status",
    colour = "ADHD Status"
  ) +
  theme(
    # bottom legend
    legend.position = "bottom",
    legend.box.margin = margin(t = -4, b = 0),
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(face = "bold", size = 7.5),
    legend.text = element_text(size = 7),
    legend.key.size = unit(0.35, "cm"),
    legend.background = element_blank()
  )
```

```{r}
#| label: statistically testing adhd wellbeing

library(rstatix)
library(dunn.test)

test_cols <- c("aff_val", "bangs_auto_frus", "bangs_comp_frus", "bangs_rel_sat", "wemwbs_total")
summary_list <- list()
kw_results   <- list()
dunn_results <- list()

for (col in test_cols) {
  cat("\n=======================================================\n")
  cat("VARIABLE:", col, "\n")
  cat("=======================================================\n")
  
  desc_summary <- mapper_full_data |>
    group_by(adhd) |>
    summarise(
      Variable = col,
      N        = sum(!is.na(.data[[col]])),
      Median   = median(.data[[col]], na.rm = TRUE),
      IQR      = IQR(.data[[col]], na.rm = TRUE),
      Mean     = mean(.data[[col]], na.rm = TRUE),
      SD       = sd(.data[[col]], na.rm = TRUE),
      .groups  = "drop"
    )
  print(desc_summary)
  summary_list[[col]] <- desc_summary
  
  form <- as.formula(paste(col, "~ adhd"))
  kw_test <- kruskal.test(form, data = mapper_full_data)
  print(kw_test)
  
  kw_eff <- mapper_full_data |> 
    kruskal_effsize(formula = form)
  print(kw_eff)
  
  kw_results[[col]] <- tibble(
    variable = col,
    chi2 = kw_test$statistic,
    df = kw_test$parameter,
    p_value = kw_test$p.value,
    effsize = kw_eff$effsize,
    magnitude = kw_eff$magnitude
  )
  
  cat("\n--- Dunn's Post-Hoc Comparisons (Holm-adjusted) ---\n")
  posthoc_dunn <- dunn.test(
    x      = mapper_full_data[[col]],
    g      = mapper_full_data$adhd,
    method = "holm",
    kw     = FALSE
  )
  dunn_results[[col]] <- posthoc_dunn
}

all_descriptives <- bind_rows(summary_list)
all_kw_tests <- bind_rows(kw_results)
```

# Latex Tables/Figures

```{r}
# descriptive statistics of intake participants
library(kableExtra)
library(modelsummary)
library(estimatr)

options(modelsummary_factory_default = "kableExtra")
options(knitr.table.format = "latex")

summary_stat_data <- intake_participants |>
  select(
    adhd,
    `Age (years)` = age,
    `Height (cm)` = height,
    `Weight (kg)` = weight,
    `Country` = country,
    `Highest Education Level` = edu_level,
    `Employment Status` = employment
  )

summary_table <- datasummary_balance(
  ~ adhd,
  data = summary_stat_data,
  fmt = 2,
  dinm = FALSE,
  output = "data.frame",
)

latex_code <- kbl(
  summary_table,
  format    = "latex",
  booktabs  = TRUE,
  linesep   = "",
  caption   = "Participant Demographic and Baseline Characteristics",
  align     = c("l", "l", rep("r", ncol(summary_table) - 2))
) |>
  add_header_above(c(
    " " = 2,
    "Diagnosed ADHD (N=331)" = 2,
    "Neurotypical (N=1475)" = 2,
    "Self-Identified ADHD (N=172)" = 2
  )) |>
  kable_styling(latex_options = c("hold_position", "scale_down"))

# writeLines(as.character(latex_code), "participant_summary_stats.tex")
```