---
title: "Analysing the final Mapper Graph"
author:
  name: Daniel Agar
format:
  pdf: default
  html:
    toc: true
  docx: default
---

```{r}
#| label: libraries
#| include: false
#| output: false
#| code-summary: "Show code (Load required R packages)"
library(tidyverse)
library(MapperAlgo)
library(igraph)
library(jsonlite)
library(knitr) # latex tables

generate_pdf_figs <- TRUE
theme_set(theme_minimal())
source("examples/open-play/helpers.R")
source("utils.R")

sourceDir("mapper", exclude = "test.R")
source("examples/open-play/mapper_data.R")
```

```{r}
#| label: data setup
#| include: false
#| cache: true
#| output: false

# Load the mapper file
grid_dir <- "examples/open-play/grid_searches"
mapper_file <- paste(
  # "shortlisted",
  # "more_telem_pca12_zscale",
  # "more_telem_pca12_final_indfix_zscale",
  "more_telem_pca12_final_zscale",
  # "more_telem_pca12_final_covextension_int10_ov40_dbscan_eps3_minPts3.json",
  "more_telem_pca12_final_covextension_int10_ov40_dbscan_eps3_minPts5.json",
  # "more_telem_pca12_final_indfix_covextension_int10_ov40_dbscan_eps0.4_minPts5.json",
  # "more_telem_pca12_covstride_int5_ov50_kmeans_max_kmeans_clusters2.json",
  # "more_telem_pca12_covstride_int10_ov50_dbscan_eps0.05_minPts3.json",
  # "more_telem_pca12_covstride_int5_ov50_dbscan_eps0.4_minPts34.json",
  sep = "/"
)
mapper_graph_data <- fromJSON(here::here(paste(grid_dir, mapper_file, sep = "/")))

any_adhd <- mapper_full_data$any_adhd
# checking data for index issues
raw_indices <- unlist(mapper_graph_data$points_in_vertex)
is_zero_indexed <- min(raw_indices, na.rm = TRUE) == 0
assigned_rows <- unique(if (is_zero_indexed) raw_indices + 1 else raw_indices)
total_rows    <- nrow(mapper_full_data)
unassigned_rows <- setdiff(seq_len(total_rows), assigned_rows)
message(sprintf(
  "Total Rows: %d | Assigned: %d | Excluded (DBSCAN Noise / Outliers): %d",
  total_rows, length(assigned_rows), length(unassigned_rows)
))
```

```{r} 
#| label: data scaling
#| output: false
scaler <- "zscale"

# Min-Max Normalisation (commonly used in Kepler Mapper)
min_max_norm <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

if (scaler == "minmax") {
  mapper_scaled <- mapper_data |>
    mutate(across(everything(), min_max_norm))
} else if (scaler == "zscale") {
  mapper_scaled <- mapper_data |>
    mutate(across(everything(), ~ scale(.)[, 1]))
} else {
  stop("No valid scaler selected")
}

mapper_pca <- prcomp(mapper_data, center = TRUE, scale = TRUE)
# print(xtable::xtable(round(cor(cbind(
#   mapper_scaled,
#   pc1 = mapper_pca$x[, 1],
#   pc2 = mapper_pca$x[, 2]
# ))[, c("pc1", "pc2")], 2),
# floating = FALSE, latex.environments = NULL, booktabs = TRUE))
```

```{r}
#| label: dimension reduction trials
#| include: false
#| eval: false

library(uwot)
library(Rtsne)

set.seed(30)
mapper_umap <- umap2(
  mapper_data,
  n_components = 2,
  scale = TRUE,
  n_neighbors = 10,
  min_dist = 0.001
)
mapper_pca <- prcomp(mapper_data, center = TRUE, scale = TRUE)
set.seed(30)
mapper_tsne <- Rtsne(
  mapper_data,
  dims = 2,
  pca_center = TRUE,
  pca_scale = TRUE
)
mapper_dists <- dist(mapper_scaled)
mapper_mds <- cmdscale(mapper_dists, k = 2)

adhd_group <- mapper_full_data$adhd

dim_red_data <- bind_rows(
  tibble(
    method = "pca",
    dim1 = mapper_pca$x[, 1],
    dim2 = mapper_pca$x[, 2],
    adhd = adhd_group
  ),
  tibble(
    method = "tsne",
    dim1 = mapper_tsne$Y[, 1],
    dim2 = mapper_tsne$Y[, 2],
    adhd = adhd_group
  ),
  tibble(
    method = "umap",
    dim1 = mapper_umap[, 1],
    dim2 = mapper_umap[, 2],
    adhd = adhd_group
  ),
  tibble(
    method = "mds",
    dim1 = mapper_mds[, 1],
    dim2 = mapper_mds[, 2],
    adhd = adhd_group
  )
)

if (generate_pdf_figs) pdf("mapper_analysis_figs/dimension_reductions_filters.pdf")
ggplot(dim_red_data, aes(x = dim1, y = dim2, colour = adhd)) +
  geom_point(size = 0.3, alpha = 0.6, shape = 1) +
  facet_wrap(vars(method), scale = "free")
if (generate_pdf_figs) dev.off()
```

```{r}
#| label: create mapper

# 10 intervals, 50% overlap
# DBSCAN(0.4, 26)
mapper <- MapperAlgo(
  original_data = mapper_scaled,
  filter_values = mapper_pca$x[, 1:2],
  percent_overlap = 40,
  methods = "dbscan",
  method_params = list(eps = 3, minPts = 5),
  cover_type = "stride",
  intervals = 10,
  num_cores = 10
)
# apply index fix
# mapper$points_in_vertex <- lapply(mapper$points_in_vertex, function(x) x - 1)
```

```{r} 
#| label: dbscan params figure

# histogram of the euclidean distances
mapper_dists <- dist(mapper_scaled)
mapper_dists <- as.numeric(mapper_dists)

if (generate_pdf_figs) pdf("mapper_analysis_figs/mapper_scaled_dists_hist.pdf")
ggplot(data.frame(distance = mapper_dists), aes(x = distance)) +
  geom_histogram(colour = "black", bins = 20) +
  scale_x_continuous(breaks = seq(0, 25, 5))
if (generate_pdf_figs) dev.off()

# List of clustering methods and their respective hyperparameter sets
# k ~ 2 x dimensions
if (generate_pdf_figs) pdf("mapper_analysis_figs/knn_distplot_zscale_k5_scaled_data.pdf")
kNNdistplot(x = mapper_scaled, k = 5) # minPts~3 is good
abline(h = 3, col = "red")
if (generate_pdf_figs) dev.off()
```
Looking at these plots and changing k in the 'kNNdistplot' gives eps~3 a good choice

```{r}
#| label: graph any adhd
if (generate_pdf_figs) pdf("mapper_analysis_figs/final_mapper_any_adhd.pdf")
create_mapper_graph(
  mapper_obj = mapper_graph_data,
  colourisation = mapper_full_data$any_adhd,
  original_data = mapper_data,
  graph_layout = layout_with_fr,
  legend = TRUE,
  legend_title = "Proportion of ADHD Participants",
  legend_tick_num = 5
)
if (generate_pdf_figs) dev.off()
```

```{r}
#| label: graph colourisation
if (generate_pdf_figs) pdf("mapper_analysis_figs/final_mapper_bangs_auto_sat.pdf")
create_mapper_graph(
  mapper_obj = mapper_graph_data,
  colourisation = mapper_data$bangs_auto_sat,
  original_data = mapper_data,
  graph_layout = layout_with_fr,
  legend = TRUE,
  legend_title = "BANGS Autonomy Satisfaction",
  legend_tick_num = 5
)
if (generate_pdf_figs) dev.off()
```

```{r}
#| label: mapper data points

# checking data for index issues
raw_indices <- unlist(mapper_graph_data$points_in_vertex)
is_zero_indexed <- min(raw_indices, na.rm = TRUE) == 0
assigned_rows <- unique(if (is_zero_indexed) raw_indices + 1 else raw_indices)
total_rows    <- nrow(mapper_full_data)
unassigned_rows <- setdiff(seq_len(total_rows), assigned_rows)
message(sprintf(
  "Total Rows: %d | Assigned: %d | Excluded (DBSCAN Noise / Outliers): %d",
  total_rows, length(assigned_rows), length(unassigned_rows)
))
noise_participants <- mapper_full_data[unassigned_rows, ]
```

## Analysing the isolated components
We shall analyse the smaller isolated component:
```{r}
#| label: isolated components

isolated_components <- find_isolated_components(mapper_graph_data, mapper_full_data)

# join node and component to the data
graph_full_data <- mapper_full_data |>
  add_column(node = 0, component = 0)

for (row in rownames(isolated_components)) {
  row_data <- isolated_components[row, ]
  node_points <- mapper_graph_data$points_in_vertex[[row_data$node]]
  graph_full_data[node_points, "node"] <- row_data$node
  graph_full_data[node_points, "component"] <- row_data$component
}


interested_cols <- c(
  "pid", "wave", "adhd",
  "wemwbs_total", "life_sat", "aff_val",
  "bangs_auto_sat", "bangs_auto_frus", "bangs_comp_sat",
   "bangs_comp_frus", "bangs_rel_sat", "bangs_rel_frus",
  "gaming_value_sum", "total_wave_minutes", "max_binge_minutes",
  "binge_day_of_week", "mean_mobile", "mean_nintendo",
  "mean_steam", "mean_xbox", "score_mean", "atten_score_mean"
)

print("Full data:")
print(skimr::skim(mapper_full_data))

# create full game data table
all_game_telemetry <- bind_rows(
  steam |>
    mutate(platform = "Steam") |>
    select(
      pid,
      platform,
      session_start = approximate_session_start,
      session_end = approximate_session_end,
      duration = minutes,
      title_id
    ),
  nintendo |>
    mutate(platform = "Nintendo") |>
    select(
      pid,
      platform,
      session_start,
      session_end,
      duration,
      title_id
    ),
  # mobile combined
  bind_rows(ios, android) |>
    mutate(platform = "Mobile", title_id = NA_character_) |>
    mutate(duration = sum(minutes), .by = c(pid, day_local)) |>
    select(
      pid,
      platform,
      session_start = day_local,
      session_end = day_local,
      duration,
      title_id
    )
) |>
  left_join(metadata, by = join_by(platform, title_id == original_name))

xbox_game_telemetry_metadata <- xbox |>
  mutate(platform = "Xbox") |>
  select(
    pid,
    platform,
    session_start,
    session_end,
    duration,
    xbox_title_id = title_id
  ) |>
  left_join(metadata, by = join_by(
    platform,
    xbox_title_id == original_name
  ))

# TODO: manual game renaming
game_sessions <- bind_rows(
  all_game_telemetry,
  xbox_game_telemetry_metadata
) |>
  mutate(
    primary_genre = case_when(
      platform == "Xbox" ~ xbox_primary_genre,
      platform == "Mobile" ~ "Mobile",
      is.na(genres) ~ NA_character_,
      TRUE ~ str_split_i(genres, ",", 1)
    ),
    secondary_genre = case_when(
      platform == "Xbox" ~ xbox_subgenre,
      platform == "Mobile" ~ "Mobile",
      is.na(genres) ~ NA_character_,
      TRUE ~ str_split_i(genres, ",", 2)
    )
  ) |>
  select(
    pid, title_id, platform,
    session_start, session_end, duration,
    starts_with("igdb_"), match_confidence,
    platforms, primary_genre, secondary_genre
  )
```

<!-- Component 1: 9 nodes 5 rows -->
```{r}
#| label: component 1

comp1_nodes <- isolated_components |>
  filter(component == 1) |>
  pull(node)
unique_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(comp1_nodes)]))
comp1_data <- mapper_full_data[unique_rows, ]
glimpse(comp1_data)

# pdf("isolated_components_pca.pdf")
# par(
#   mar = c(3.2, 3.2, 0.8, 0.8), # bottom, left, top, right
#   mgp = c(1.8, 0.5, 0), # distance for c(axis title, tick labels, tick line)
#   tcl = -0.3 # slightly shorter tick marks
# )
plot(
  mapper_pca$x[, 1:2],
  col = "black",
  pch = 16,
  cex = 0.6,
  main = NULL,
  xlab = "PC1",
  ylab = "PC2"
)
points(
  mapper_pca$x[unique_rows, 1:2, drop = FALSE],
  col = "red",
  pch = 19,
  cex = 1.2
)
# dev.off()
```
Five rows of data, 3 from one participant.
Low affective valence (0, 5, 50 being the one highest) and wellbeing (wemwbs 7-11).
Very high BANGS autonomy frustration.
they felt forced into actions and choices, most of which they found boring.
They always had max scores (21) for BANGS relatedness frustration and very low satisfaction.
They clearly didn't feel they felt cared for in the games or had connections.
Their interactions made them feel severely unwelcome and were toxic towards them.

The games they played need investigating.
```{r} 
#| label: checking comp 1 games

comp1_pid_info <- comp1_data |>
  select(pid, wave, linked_platforms) |>
  left_join(wave_daily_telemetry, by = join_by(pid, wave)) |>
  select(
    pid, wave, linked_platforms,
    wave_start_date = min_twoweek_date,
    wave_end_date = survey_date
  ) |>
  distinct()

# get dates valid for the wave
comp1_games <- comp1_pid_info |>
  # steam
  left_join(game_sessions, by = join_by(
    pid,
    wave_start_date <= session_start,
    wave_end_date > session_end
  )) |>
  mutate(
    minutes_played = sum(duration),
    .by = c(pid, title_id)
  ) |>
  select(-duration) |>
  distinct(pid, title_id, .keep_all = TRUE)
```

Game data reveals p3011712557 (neurotypical) had the largest of 2693.7 minutes played in wave 3.
They played 9 days in this wave with very high gaming influence on social life.
Max bangs rel frus, comp frus and auto frus.
They played World and Tanks and Counter Strike 2 during the wave, which explains the high frustrations as they are competitive shooters.
Participant p1219663417 had 3 waves of data (diagnosed ADHD.)
Wave 1 showed they just played portal 2 (puzzle game) for 29 minutes across 2 days.
BANGS rel frus (21), auto frus (19) and low satisfaction across highlights.
Puzzle games are typically mentally-challenging and single player, so poor relatedness is expected.
Poor autonomy is also expected -- a puzzle game won't let you continue till you complete the level.
In wave 3 they played for 470 minutes across five days, playing role playing and beat-em-up games.
Still believed they couldn't play as they wanted.
they played for 2700 in wave 5 (CS2=1132 high binge 810, Hundred line=386, multiplayer casual total 280, elden ring 135).
Participant p7842684564 played 1504 with binge 430.
Mostly in Souce SDK (1241) and then Schedule 1 (144).
Source SDK 2007 is a software development kit for creating mods for games on source engine.
This is not a game as such, rather development/work.
To summarise: the isolated component consists of participants with negative feelings towards the video games they have been playing, along with very low wemwbs and affective valence scores (outliers).

## Main Structure
```{r}
#| label: main structure description

main_nodes <- isolated_components |>
  filter(component == 2) |>
  pull(node)

unique_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(main_nodes)]))
main_data <- mapper_full_data[unique_rows, ]
skimr::skim(main_data)
```

The main structure consists of 72 nodes and more than 900 rows of data.
There are small tails on the end corners.
There appears to be an increase in ADHD proportions at the corners.
As demonstrated through the PCA correlations, we see positive experiences and wellbeing decrease as the the node size decreases.



## Old version: 30% overlap and less columns
<!-- Component 2: 4 nodes 1 row -->
```{r}
#| label: component 2

comp2_nodes <- isolated_components |>
  filter(component == 2) |>
  pull(node)
unique_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(comp2_nodes)]))
comp2_data <- mapper_full_data[unique_rows, ]
glimpse(comp2_data)

plot(
  mapper_pca$x[, 1:2],
  col = "black",
  pch = 16,
  cex = 0.6,
  main = "Component 6 data highlight",
  xlab = "PC1",
  ylab = "PC2"
)
points(
  mapper_pca$x[unique_rows, 1:2, drop = FALSE],
  col = "red",
  pch = 19,
  cex = 1.2
)
```

The isolated component of 4 nodes consists of only one data point from the biweekly data.
They are neurotypical, had their highest binge of wave 1 on a Tuesday.
They played very little during the two weeks (234 minutes).
As the total wemwbs score for someone saying "some of the time" to the positive questions is 22, they appear to have slightly lower wellbeing, but not noticeably.
However, they say that gaming has an influce on their life (gaming_value_sum = 26) and is almost classed as an outlier based on the 1.5IQR rule.
Switching to the gaming value colourisation shows this as red.
They also show higher BANGS relatedness satisfaction (17), which is significantly higher than average.
Despite not playing lots during the time, half their wave telemetry was in one session, and the participant felt a strong connection with the players/characters around them.
As they play steam games, I want to check the metadata for this player, looking into what they generally play...

<!-- Component 3: 1 node 1 point -->
```{r}
#| label: component 3

comp3_nodes <- isolated_components |>
  filter(component == 3) |>
  pull(node)
unique_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(comp3_nodes)]))
comp3_data <- mapper_full_data[unique_rows, ]
glimpse(comp3_data)
plot(
  mapper_pca$x[, 1:2],
  col = "black",
  pch = 16,
  cex = 0.6,
  main = "Component 3 data highlight",
  xlab = "PC1",
  ylab = "PC2"
)
points(
  mapper_pca$x[unique_rows, 1:2, drop = FALSE],
  col = "red",
  pch = 19,
  cex = 1.2
)
```

Another component is a singular node and data point.
They are a very interesting case:
low wemwbs (12), life satisfaction (2), affective valence (40), very high autonomy frustration (20), high competence satisfaction (20), very high relatedness frustration (21, the maximum value).
They played a total of more than 1740 minutes in the two weeks (29 hours), with over 500 minutes in one session.
Although only on instance and is not actually a high total telemetry, the long session highlights extreme hyperfocus.
They also achieved a higher attention score of 68.
The video game type needs to be checked; playing first person shooters may show higher relatedness frustration as players can experience toxic behaviour.

```{r} 
#| label: checking comp 3 games

comp3_pid_info <- comp5_data |>
  select(pid, wave, linked_platforms) |>
  left_join(wave_daily_telemetry, by = join_by(pid, wave)) |>
  select(pid, wave, linked_platforms, wave_start_date = min_twoweek_date, wave_end_date = survey_date) |>
  distinct()

# get dates valid for the wave
comp3_games <- steam |>
  right_join(comp3_pid_info, by = join_by(
    pid,
    approximate_session_start >= wave_start_date,
    approximate_session_start < wave_end_date
  )) |>
  group_by(title_id) |>
  summarise(minutes_played = sum(minutes)) |>
  ungroup() |>
  left_join(metadata, by = join_by(title_id == original_name))
glimpse(comp3_games)
```
This player spent all but one minute of their play time playing "Vampire Surivors", a role-playing survival game hunting vampires.

<!-- Component 5: 1 node 1 row -->
```{r}
#| label: component 5

comp5_nodes <- isolated_components |>
  filter(component == 5) |>
  pull(node)
unique_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(comp5_nodes)]))
comp5_data <- mapper_full_data[unique_rows, ]
glimpse(comp5_data)
plot(
  mapper_pca$x[, 1:2],
  col = "black",
  pch = 16,
  cex = 0.6,
  main = "Component 5 data highlight",
  xlab = "PC1",
  ylab = "PC2"
)
points(
  mapper_pca$x[unique_rows, 1:2, drop = FALSE],
  col = "red",
  pch = 19,
  cex = 1.2
)
```

There is another singular node and data point.
They are a neurotypical participant and with their total telemetry (687) almost being their highest binge session (686) on a saturday.
This also exhibits hyperfocus, an ADHD trait.
But with very few sessions it does not mean they secretly have ADHD.
They are the 75th percentile in wemwbs (high wellbeing) but have a very low affective valence (20) and life satisfaction (2).
This is odd, and could be an input error.
Gaming does not have a large influence on their social life.
They experience very low satisfactions (3, 3, 3 for all) but also lower frustrations (9, 9 for comp rel).
Interestingly, they have exceedingly high autonomy frustration.
With such a large frustration, why do they have a large binge session of more than 11 hours?
I need to check the game they played, and why the inputs to the survey are odd (just filling it out weirdly, although they didn't fail the check).

<!-- Component 4: 3 nodes 2 rows -->
```{r}
#| label: component 4

comp4_nodes <- isolated_components |>
  filter(component == 4) |>
  pull(node)
unique_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(comp4_nodes)]))
comp4_data <- mapper_full_data[unique_rows, ]
glimpse(comp4_data)
plot(
  mapper_pca$x[, 1:2],
  col = "black",
  pch = 16,
  cex = 0.6,
  main = "Component 4 data highlight",
  xlab = "PC1",
  ylab = "PC2"
)
points(
  mapper_pca$x[unique_rows, 1:2, drop = FALSE],
  col = "red",
  pch = 19,
  cex = 1.2
)
```

An isolated component with 3 nodes and 2 rows of data (from two different participants) contains one diagnosed ADHD participant.
They are also from two different waves (2 and 3 diagnosed) and have differing wellbeings (20, 10 diagnosed).
Diagnosed experienced extreme autonomy frustration, and very low satisfaction, but played a lot less than the neurotypical (477 total compared to 5007).
However, their life satisfaction is low (3, 0 diagnosed) and their affective valence (40, 5 diagnosed).

<!-- Component 6: 2 nodes 3 rows -->
```{r}
#| label: component 6
comp6_nodes <- isolated_components |>
  filter(component == 6) |>
  pull(node)
unique_rows <- unique(unlist(mapper_graph_data$points_in_vertex[c(comp6_nodes)]))
comp6_data <- mapper_full_data[unique_rows, ]
glimpse(comp6_data)
plot(
  mapper_pca$x[, 1:2],
  col = "black",
  pch = 16,
  cex = 0.6,
  main = "Component 6 data highlight",
  xlab = "PC1",
  ylab = "PC2"
)
points(
  mapper_pca$x[unique_rows, 1:2, drop = FALSE],
  col = "red",
  pch = 19,
  cex = 1.2
)
```
Another component contains just 2 nodes but three participants (one is self-identified ADHD).
Two are from the second wave, and one from the sixth.
They have varied affective valence (40, 59, 80) and BANGS.
However, they have consistent attention scores (the self-identified participant has the attention score in the middle).
Two of the total wave minutes are similar (1830, 1787) but one is significantl lower (363).
I cannot find a pattern between the three data points, so I wonder if this is the limitation of using PCA with limited variance.
Looking at the rotations, I don't see any reason for these data points to be clustered together.
