---
title: "Utilising Mapper with Open Play"
author:
  name: Daniel Agar
format:
  pdf: default
  html:
    toc: true
  docx: default
---

```{r}
#| label: libraries
#| include: false
#| code-summary: "Show code (Load required R packages)"
library(tidyverse)
library(data.table)
library(dtplyr)
library(skimr)
library(stringr)
library(MapperAlgo)
library(jsonlite)
library(dbscan)
library(patchwork)
library(tidymodels)
# library(uwot)
library(rpart.plot)
library(probably)
library(Rtsne)
library(GGally)
library(isotree)
library(MLmetrics)

theme_set(theme_minimal())
source("examples/open-play/helpers.R")
source("utils.R")

sourceDir("mapper", exclude = "test.R")
```

```{r}
#| label: data setup
#| include: false
#| cache: true
#| output: false

android <- read_csv("data/open-play/clean/android.csv.gz")
metadata <- read_csv("data/open-play/clean/game_metadata.csv.gz")
ios <- read_csv("data/open-play/clean/ios.csv.gz")
nintendo <- read_csv("data/open-play/clean/nintendo.csv.gz")
simon <- read_csv("data/open-play/clean/simon.csv.gz")
steam <- read_csv("data/open-play/clean/steam.csv.gz")
steam_visibility <- read_csv(
  "data/open-play/raw/telemetry_steam_account_linking_raw.csv.gz"
)
biweekly <- read_csv("data/open-play/clean/survey_biweekly_final.csv.gz")
daily <- read_csv("data/open-play/clean/survey_daily_final.csv.gz", guess_max = 10000)
intake_participants <- read_csv("data/open-play/clean/survey_intake_participants_final.csv.gz", guess_max = 34000)
participant_coords <- read_csv("data/open-play/clean/participant_coords.csv.gz")
time_use <- read_csv("data/open-play/clean/timeuse.csv.gz") |>
  filter(quality_flag == "Good quality")
xbox <- read_csv("data/open-play/clean/xbox.csv.gz")
```

# Getting the Data Mapper Ready
I need to bare in mind that for this to work with the telemetry data,
I should either cut the telemetry data prior to surveys,
or have a variable to indicate how much they played on average beforehand.

Any transformation possibly needed:
- pid: post-mapper analysis
- demographic variables: require some form of encoding/scaling
- day of study: this will be useful for a temporal tracking (true dates are irrattic)
- location: convert country and geo_area into coordinates

```{r}
#| label: data-prep
#| cache: true
#| include: false

# turning country and geo code into long/lat for mapper
# these are the centroids of the combined addresses
# participant_coords <- intake_participants |>
#   select(country, geo_area) |>
#   distinct() |>
#   mutate(address = paste(geo_area, country, sep = ", ")) |>
#   geocode(address = address) |>
#   select(-address)

# write_csv(participant_coords, file = "data/open-play/clean/participant_coords.csv.gz")

# merge with the local time zone offsets from intake participants
# Note: We keep country and local_timezone for DST-aware conversion
tz_map <- intake_participants |>
  mutate(
    pid,
    country,
    local_timezone,
    .keep = "none"
  ) |>
  distinct(pid, .keep_all = TRUE)

# aggregate data at each level of granularity (session, hourly, daily)
# session level (Nintendo, Xbox, Steam already at session level)
session_telemetry <- bind_rows(
  xbox |> mutate(platform = "Xbox"),
  nintendo |> mutate(platform = "Nintendo"),
) |>
  left_join(tz_map, by = "pid") |>
  filter(!is.na(local_timezone)) |>
  mutate(
    # Calculate DST-aware offset for each timestamp
    offset_start = get_dst_offset(session_start, country, local_timezone),
    offset_end = get_dst_offset(session_end, country, local_timezone),
    # Add offset to get local time values (keeping UTC label for compatibility)
    start_local = session_start + offset_start,
    end_local = session_end + offset_end,
    duration_min = as.numeric(difftime(
      session_end,
      session_start,
      units = "mins"
    ))
  ) |>
  filter(
    !is.na(session_start),
    !is.na(session_end),
    session_end > session_start,
    duration_min >= 1
  )

# Hourly from sessions (Nintendo + Xbox)
# Join with timezone map and calculate DST-aware offsets
# Note: start_local/end_local have UTC labels but local time VALUES (for hour extraction)
sessions_telemetry <- bind_rows(
  xbox |> mutate(platform = "Xbox"),
  nintendo |> mutate(platform = "Nintendo")
) |>
  mutate(pid = as.character(pid)) |>
  left_join(tz_map, by = "pid") |>
  filter(!is.na(local_timezone)) |>
  mutate(
    # Calculate DST-aware offset for each timestamp
    offset_start = get_dst_offset(session_start, country, local_timezone),
    offset_end = get_dst_offset(session_end, country, local_timezone),
    # Add offset to get local time values (keeping UTC label for compatibility)
    start_local = session_start + offset_start,
    end_local = session_end + offset_end,
    duration_min = as.numeric(difftime(
      session_end,
      session_start,
      units = "mins"
    ))
  ) |>
  filter(
    !is.na(session_start),
    !is.na(session_end),
    session_end > session_start,
    duration_min >= 1
  )

hourly_from_sessions <- sessions_telemetry |>
  filter(!is.na(start_local), !is.na(end_local)) |>
  mutate(
    h0_local = floor_date(start_local, "hour"),
    h1_local = floor_date(end_local - seconds(1), "hour"),
    n_hours = as.integer(difftime(h1_local, h0_local, units = "hours")) + 1
  ) |>
  filter(!is.na(n_hours), n_hours > 0) |>
  tidyr::uncount(n_hours, .remove = FALSE, .id = "k") |>
  mutate(
    hour_start_local = h0_local + hours(k - 1),
    minutes = pmax(
      0,
      as.numeric(difftime(
        pmin(end_local, hour_start_local + hours(1)),
        pmax(start_local, hour_start_local),
        units = "mins"
      ))
    ),
    # Convert back to UTC (this preserves the instant, just changes label)
    hour_start_utc = with_tz(hour_start_local, tzone = "UTC")
  ) |>
  select(pid, platform, hour_start_local, hour_start_utc, minutes)

# --- Hourly from Steam (already hourly) -------------------------------------
hourly_from_steam <- steam |>
  select(pid, datetime_hour_start, minutes) |>
  mutate(pid = as.character(pid)) |>
  left_join(tz_map, by = "pid") |>
  filter(!is.na(local_timezone)) |>
  mutate(
    platform = "Steam",
    hour_start_utc = datetime_hour_start,
    # Calculate DST-aware offset and add to get local time values
    offset = get_dst_offset(datetime_hour_start, country, local_timezone),
    hour_start_local = datetime_hour_start + offset
  ) |>
  select(pid, platform, hour_start_local, hour_start_utc, minutes)

hourly_telemetry <- bind_rows(hourly_from_sessions, hourly_from_steam)

# Daily (Nintendo + Xbox + Steam; collapse hourly to days)
daily_telemetry <- hourly_telemetry |>
  mutate(
    day_local = as.Date(hour_start_local),
  ) |>
  group_by(pid, platform, day_local) |>
  summarise(minutes = sum(minutes, na.rm = TRUE), .groups = "drop") |>
  bind_rows(ios) |>
  bind_rows(android)

daily_ios_android <- bind_rows(
  ios |>
    transmute(
      pid = as.character(pid),
      platform = "iOS",
      day_local = as.Date(day_local),
      minutes
    ),
  android |>
    transmute(
      pid = as.character(pid),
      platform = "Android",
      day_local = as.Date(day_local),
      minutes
    )
)

daily_all <- bind_rows(daily_telemetry, daily_ios_android) |>
  group_by(pid, day_local) |>
  summarise(total_minutes = sum(minutes, na.rm = TRUE), .groups = "drop")

# Weekly
weekly_all <- daily_telemetry |>
  mutate(
    week = floor_date(day_local, "week")
  ) |>
  group_by(pid, platform, week) |>
  summarise(minutes = sum(minutes, na.rm = TRUE), .groups = "drop")

telemetry_spans <- daily_telemetry |>
  group_by(pid, platform) |>
  summarise(
    telemetry_start = min(day_local, na.rm = TRUE),
    # end of last hour bin
    telemetry_end = max(day_local, na.rm = TRUE) + hours(1),
    week = floor_date(telemetry_end, "week"),
    n_weeks = as.integer(difftime(
      telemetry_end,
      telemetry_start,
      units = "weeks"
    )) +
      1,
    .groups = "drop"
  )
```

```{r}
demographic_vars <- c(
  "pid",
  "country",
  "geo_area",
  "gender",
  "height",
  "weight",
  "age",
  "ethnicity",
  "edu_level",
  "employment",
  "political_party",
  "marital_status",
  "adhd"
)
```

# Mapper Tests

Here are some mapper configurations I want to run:
- Include categorical data with Gower Distance
- Telemetry data only (study date, play time per platform)
- Well-being data only
- Telemetry and well-being data

## Telemetry data
For this experiment, I will use the telemetry data as the basis of the filter function.
The reason for this is that we can explore gaming patterns to find any clusters/groupings (ADHD related telemetry).
This can be followed up with well-being overlays.

Based on literature, I will try with PCA.
We will have a row for each day of the survey for each participant.
Each row will have: minutes played for each platform, day of the study.

```{r}
# creates exhaustive entry list for all 12 weeks from the start date
# if there are no entries for a date, create a new row with 0 for each platform playtime

platforms <- c("Steam", "Nintendo", "Xbox", "Android", "iOS")

# study lasts 12 weeks
# start at 0 for first day
study_duration_days <- 83

daily_start_dates <- daily |>
  group_by(pid) |>
  mutate(first_wave = min(wave)) |>
  ungroup() |>
  filter(wave == first_wave) |>
  select(pid, daily_start_date = date, first_wave)

biweekly_start_dates <- biweekly |>
  group_by(pid) |>
  mutate(first_wave = min(wave)) |>
  ungroup() |>
  filter(wave == first_wave) |>
  select(pid, biweekly_start_date = date, first_wave)

survey_start_dates <- biweekly_start_dates |>
  full_join(daily_start_dates, by = join_by(pid)) |>
  mutate(started_study = as.Date(pmin(biweekly_start_date, daily_start_date, na.rm = TRUE))) |>
  select(pid, started_study)

last_daily_survey <- daily |>
  group_by(pid) |>
  summarise(last_daily_survey = max(date, na.rm = TRUE))

last_biweekly_survey <- biweekly |>
  group_by(pid) |>
  summarise(last_biweekly_survey = max(date, na.rm = TRUE))

last_daily_telemetry <- daily_telemetry |>
  group_by(pid) |>
  summarise(last_ping = max(day_local, na.rm = TRUE))

participant_windows <- survey_start_dates |>
  left_join(last_daily_survey, by = "pid") |>
  left_join(last_biweekly_survey, by = "pid") |>
  left_join(last_daily_telemetry, by = "pid") |>
  mutate(
    last_survey_date = pmax(last_daily_survey, last_biweekly_survey, na.rm = TRUE),
    active_study_days = as.numeric(difftime(last_survey_date, started_study, units = "days")) + 1,
    active_study_days = pmin(active_study_days, study_duration_days)
  )

# participant timelines dynamically creating rows up till the last day of participation
study_timeline <- participant_windows |>
  filter(!is.na(active_study_days)) |> # drop pid with zero usable dates
  mutate(study_day_offset = map(active_study_days, ~0:(.x - 1))) |>
  unnest(study_day_offset) |>
  mutate(day_local = started_study + study_day_offset) |>
  select(pid, started_study, last_survey_date, active_study_days, day_local, study_day_offset)

wide_daily_telemetry <- daily_telemetry |>
  distinct(pid, platform, day_local, .keep_all = TRUE) |>
  pivot_wider(
    names_from = platform,
    names_glue = "{platform}_minutes",
    values_from = minutes,
    values_fill = 0
  ) |>
  rename_with(tolower)

complete_daily_telemetry <- study_timeline |>
  left_join(wide_daily_telemetry, by = join_by(pid, day_local)) |>
  mutate(
    across(ends_with("_minutes"), ~replace_na(.x, 0)),
    day_of_week = wday(day_local)
  )

telemetry_traits <- complete_daily_telemetry |>
  group_by(pid, active_study_days) |>
  summarise(
    # activity stats
    mean_daily_steam_minutes = sum(steam_minutes, na.rm = TRUE) /
      first(active_study_days),
    mean_daily_nintendo_minutes = sum(nintendo_minutes, na.rm = TRUE) /
      first(active_study_days),
    mean_daily_xbox_minutes = sum(xbox_minutes, na.rm = TRUE) /
      first(active_study_days),
    mean_daily_mobile_minutes = sum(ios_minutes + android_minutes, na.rm = TRUE) /
      first(active_study_days),
    mean_daily_total = sum(
      steam_minutes + nintendo_minutes + xbox_minutes +
        ios_minutes + android_minutes, na.rm = TRUE
    ) / first(active_study_days),

    max_binge_day = max(steam_minutes + nintendo_minutes + xbox_minutes + ios_minutes + android_minutes, na.rm = TRUE),

    # gaming traits
    active_gaming_ratio = mean((steam_minutes + nintendo_minutes + xbox_minutes + ios_minutes + android_minutes) > 0),
    .groups = "drop_last"
  ) |>
  ungroup()

interested_deomographic_cols <- c(
  "adhd",
  "country", "long", "lat",
  "age", "gender", "ethnicity", "height", "weight",
  "edu_level", "employment", "marital_status"
)
telem_traits_dem <- intake_participants |>
  left_join(participant_coords, by = join_by(country, geo_area)) |>
  select(all_of(c("pid", interested_deomographic_cols))) |>
  right_join(telemetry_traits, by = "pid") |>
  drop_na()

# old: add to mapper data
tel_dem_data_full <- intake_participants |>
  left_join(participant_coords, by = join_by(country, geo_area)) |>
  left_join(complete_daily_telemetry, by = join_by(pid))
```

### Old: combinations of telemetry
I will choose the variables.
Given that there are lots of variables, I cannot just use an identity filter function.
Therefore, PCA may be more appropriate here
```{r}
continuous_demographic_cols <- c("age", "height", "weight", "long", "lat")
categoric_game_cols <- c("plays_xbox", "plays_steam", "plays_nintendo", "plays_ios", "plays_android")
study_cols <- c("study_day_offset")
telemetry_cols  <- c(
  "steam_minutes",
  "nintendo_minutes",
  "xbox_minutes",
  "ios_minutes",
  "android_minutes"
)

tel_dem_data_mapper <- tel_dem_data_full |>
  select(all_of(c(continuous_demographic_cols, study_cols, telemetry_cols)))

# filter function: first two principal components
tel_dem_pca <- prcomp(tel_dem_data_mapper, scale. = TRUE)
tel_dem_filter_fn <- tel_dem_pca$rotation[, 1:2]
summary(tel_dem_pca)

# ggplot(tel_dem_pca$x, aes(x = PC1, y = PC2, colour = tel_dem_data_full$adhd)) +
#   geom_point()
```
PCA fails to capture enough of the variance to warrant its use as a filter function.
I will try UMAP:
```{r}
set.seed(30)
# tel_dem_umap <- umap2(
#   tel_dem_data_mapper,
#   n_components = 2,
#   scale = TRUE,
#   n_neighbors = 15, # default
#   min_dist = 0.1 # default
# )
```

I found we had a lot of days not being played (found in the distributions).
Instead, I will just do descriptive statistics for each participant during the study.

### Participant-based Telemetry
Here, due to the small amount of data when summarised, I will trial different filter functions.

#### PCA

```{r}
n_components <- 2

# data scaling
telem_traits_scaled <- telem_traits_dem |>
  select(starts_with("mean_daily"), active_gaming_ratio, max_binge_day) |>
  # scale by dividing by standard deviation
  mutate(across(everything(), ~scale(.)[,1]))

telem_traits_pca <- prcomp(telem_traits_scaled)
summary(telem_traits_pca)

telem_trailts_dist <- dist(telem_traits_scaled, method = "euclidean")
telem_traits_filter <- telem_traits_pca$x[, 1:n_components]
```

##### Mapper Parameter Grid Search
```{r}
telem_traits_mapper_search_time <- system.time({
  telem_traits_mapper_search <- GridSearch(
    original_data = telem_traits_scaled,
    filter_values = telem_traits_filter,
    label = telem_traits_dem$adhd,
    column = "ADHD",
    cover_type = "stride",
    width_vec = c(0.5, 1, 1.5),
    overlap_vec = c(50),
    num_cores = 12,
    out_dir = "examples/open-play/telem_traits_mapper_grid"
  )
})
```
Now I need to find good values for DBSCAN.
The rule of thumb for `minPts` is either 2x the dimensions, or +1.
For `eps`, a k-distance graph showing the distance to each k-th neighbour for every point.
With this, I am looking for a value near the 'elbow'.

```{r}
# rules of thumb for dbscan:
# minPts: >= d + 1, 2 x d
# eps: kNN distance graph to find elbow
k_val <- ncol(telem_traits_scaled) + 1
kNNdistplot(telem_traits_scaled, k = k_val)
```
This shows a value between 1.5 and 1.8 will split majority of the data and isolate the outliers.
For a more localised graph, I could use a smaller value (~1/1.2).

```{r}
# best so far was 20 intervals, 1.4 eps
telem_traits_mapper <- MapperAlgo(
  original_data = telem_traits_scaled,
  filter_values = telem_traits_filter,
  percent_overlap = 50,
  methods = "dbscan",
  method_params = list(eps = 1.5, minPts = 9),
  cover_type = "stride",
  intervals = 10,
  num_cores = 12
)

# save the file in the 'mappers' directory
export_data <- list(
  adjacency = telem_traits_mapper$adjacency,
  num_vertices = telem_traits_mapper$num_vertices,
  level_of_vertex = telem_traits_mapper$level_of_vertex,
  points_in_vertex = telem_traits_mapper$points_in_vertex,
  original_data = telem_traits_dem
)

mapper_trials_dir <- file.path("examples", "open-play", "mappers")
if (!dir.exists(mapper_trials_dir)) {
  dir.create(mapper_trials_dir)
}

file.create(file.path(mapper_trials_dir, "telemetry_traits.json"))
write(
  toJSON(export_data, auto_unbox = TRUE),
  file = here::here(file.path(mapper_trials_dir, "telemetry_traits.json"))
)
```

```{r}
MapperPlotter(
  telem_traits_mapper,
  telem_traits_dem,
  label = telem_traits_dem$adhd,
  avg = FALSE
)

create_mapper_graph(
  telem_traits_mapper,
  colourisation = telem_traits_dem$adhd,
  original_data = telem_traits_dem
)
```
Before I fully analyse it, I need to 
This mapper graph shows a small grouping of self-diagnosed participants, which is interesting.
There is also a tail in mapper graph which shows just the ADHD groups.
This was without the active_study_days (enabling some temporal reasoning).
Although arguably, including the `active_study_days` may not change their telemetry data -
you can play video games without doing the well-being survey.

Nodes 2, 3, and 6 show the tail of Diagnosed ADHD participants, with 4, 9, and 5 being Self-Identified ADHD participants.
```{r}
mapper_node2 <- get_node_details(telem_traits_mapper, telem_traits_dem, 2)
mapper_node2 |>
  mutate(adhd = as.factor(adhd)) |>
  skim()
print(mapper_node2)
```
Node 2 consists of a sole american woman with diagnosed ADHD.
From this participant, they participated in 44.7 days of the study.
They only play on nintendo and mobile games, have a mean nintendo daily playtime of 196 minutes.
However, they played 673 minutes in one day, which is a very large amount.

```{r}
mapper_node3 <- get_node_details(telem_traits_mapper, telem_traits_dem, 3)
mapper_node3 |>
  mutate(adhd = as.factor(adhd)) |>
  skim()
print(mapper_node3)
```
The same as node 2.

```{r}
mapper_node6 <- get_node_details(telem_traits_mapper, telem_traits_dem, 6)
mapper_node6 |>
  mutate(adhd = as.factor(adhd)) |>
  summary()
print(mapper_node6)
```
The same as node 2 again.
It appears the 50 percent overlap and 10 intervals is finding inviduals who are very much outside of the population average.

In conclusion, this mapper graph is actually showing very little

```{r}
mapper_node4 <- get_node_details(telem_traits_mapper, telem_traits_dem, 4)
mapper_node4 |>
  mutate(adhd = as.factor(adhd)) |>
  summary()
print(mapper_node4, width = Inf)
```
This is another american woman, but is self-diagnosed ADHD.
They have participated in only around 3 days of the study, but also have a large binge day and similar daily total to the previous participant.
It seems that the 'ADHD gold-nugget' found in the mapper graph is actually only identified 2 of the ADHD-related participants from the 502 in the data.

```{r}
mapper_node9 <- get_node_details(telem_traits_mapper, telem_traits_dem, 9)
mapper_node9 |>
  mutate(adhd = as.factor(adhd)) |>
  summary()
print(mapper_node9, width = Inf)
```
The same woman.

```{r}
mapper_node5 <- get_node_details(telem_traits_mapper, telem_traits_dem, 5)
mapper_node5 |>
  mutate(adhd = as.factor(adhd)) |>
  summary()
print(mapper_node5, width = Inf)
```
This node clearly shows the joining of two participants.

The flare is characterised by the large nintendo playtimes, which also happens to be two participants that have ADHD.

## Retrying with More columns
I feel that just the summarised telemetry was a mistake.
Mapper is a exploratory method, and limiting it to such summarisations hinders its unique ability for visual exploration.
Here, I will try to include as much telemetry and well-being data.
Specifically, I will now try to have a row per survey response, ideally joined by telemetry matching the survey timeframe,
and any variables that have relation to ADHD.

The new perspective of this mapper graph is to have one lens focussing on a way to distinguish between ADHD types,
and another lens to spread the data so we can pull more information from it.

### Data Preparation
The gdt variables are mainly answered at the start of the study (wave 1) and at the end (wave 3).
The self reported variables have a lot of missingness and shouldn't be used if I want lots of data for mapper.

This cleaned biweekly survey data leaves us with lots of rows to analyse (6294).

#### Telemetry Columns
Now I need to join on the telemetry data that matches.
This will be detailed descriptive statistics about the telemetry.
This includes: average session time for each day of the week,
maybe hourly metrics too?
average session time per platform,
standard deviation in these session metrics,
longest session,
day of longest session (during the week),
average number of sessions a day,
latest time to start a session during the working week,
latest time to start a session at the weekend,
earliest time to start a session during the working week,
earliest time to start a session at the weekend,

```{r}
# average session time for each day of the week

# join daily telemetry up to two weeks prior to the biweekly survey
# excluding the lower bound, to include the current day of the survey
wave_daily_telemetry <- biweekly |>
  left_join(select(intake_participants, pid, adhd), by = "pid") |>
  # cuts a lot of bangs missingness
  filter(played_any_games == "Yes") |>
  select(pid, survey_date = date, wave) |>
  drop_na() |>
  mutate(
    survey_date = as.Date(survey_date),
    min_twoweek_date = survey_date - weeks(2)
  ) |>
  left_join(
    daily_telemetry,
    by = join_by(
      pid,
      # get daily telemetry up to two weeks before response
      # excludes survey day (could play more after survey)
      min_twoweek_date <= day_local,
      survey_date > day_local
    ),
    # account for response delays impacting other responses
    relationship = "many-to-many"
  ) |>
  # start of the week will be Monday
  # merge mobile instances
  mutate(
    day_of_week = lubridate::wday(day_local, week_start = 1),
    platform = case_when(
      platform == "iOS" ~ "Mobile",
      platform == "Android" ~ "Mobile",
      TRUE ~ platform
    )
  ) |>
  # handle the duplicated "Mobile" rows
  group_by(across(c(-minutes))) |>
  summarise(minutes = sum(minutes), .groups = "drop") |>
  # complete cases (0 for missing days)
  group_by(pid, wave) |>
  complete(
    day_local = seq.Date(
      first(min_twoweek_date),
      first(survey_date),
      by = "day"
    ),
    survey_date = first(survey_date),
    min_twoweek_date = first(min_twoweek_date),
    platform = c("Mobile", "Steam", "Xbox", "Nintendo"),
    fill = list(minutes = 0)
  ) |>
  ungroup() |>
  mutate(day_of_week = lubridate::wday(day_local, week_start = 1)) |>
  filter(!is.na(day_local) & !is.na(platform))

wave_daily_totals <- wave_daily_telemetry |>
  pivot_wider(
    names_from = platform,
    values_from = minutes,
    names_glue = "{platform}_{.value}",
  ) |>
  rename_with(tolower) |>
  mutate(
    total_minutes = steam_minutes + xbox_minutes +
      mobile_minutes + nintendo_minutes
  )

biweekly_mapper_telemetry <- wave_daily_totals |>
  group_by(pid, wave) |>
  summarise(
    # telemetry frequency
    days_played = sum(total_minutes > 0),
    total_wave_minutes = sum(total_minutes),
    # central tendendencies and spread of telemetry
    mean_daily_minutes = mean(total_minutes),
    median_daily_minutes = median(total_minutes),
    sd_daily_minutes = sd(total_minutes),
    # coefficient of variation of daily minutes
    cv_daily_minutes = if_else(
      mean_daily_minutes == 0,
      0,
      sd_daily_minutes / mean_daily_minutes
    ),
    # hyperfocus/binge metrics
    max_binge_minutes = max(total_minutes),
    binge_day_of_week = if_else(
      max_binge_minutes > 0,
      day_of_week[which.max(total_minutes)],
      NA_real_ # unlikely to happen, but check
    ),
    # structure of telemetry metrics
    mean_weekday_minutes = mean(total_minutes[day_of_week %in% 1:5]),
    mean_weekend_minutes = mean(total_minutes[day_of_week %in% 6:7]),
    # context switching
    mean_mobile = mean(mobile_minutes),
    mean_steam = mean(steam_minutes),
    mean_nintendo = mean(nintendo_minutes),
    mean_xbox = mean(xbox_minutes),
    platforms_used = sum(c(
      any(mobile_minutes > 0),
      any(steam_minutes > 0),
      any(xbox_minutes > 0),
      any(nintendo_minutes > 0)
    )),
    binge_day_of_week = factor(
      binge_day_of_week,
      levels = c(1, 2, 3, 4, 5, 6, 7),
      labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
    ),
    binge_day_of_week = fct_na_value_to_level(binge_day_of_week, level = "None"),
    # keep pid and wave columns
    .groups = "drop"
  )
```
Now I have my biweekly telemetry data!

However, I need to sort my cognitive data.
As there is only 18% of the rows with complete data (all waves), I cannot simply use a long format.
Therefore, I can just use a average of the values.
This will guarantee a single metric, but loses the temporal changes that I would like to see through ADHD participants.
To overcome this, I could plot regression models of the metrics over the waves, and use the gradient as a new feature.
That way, single responses report no change, and multiple responses show the changes through time.
```{r}
simon |>
  left_join(select(intake_participants, pid, adhd)) |>
  group_by(adhd, pid) |>
  summarise(waves_completed = n_distinct(wave), .groups = "drop") |>
  count(adhd, waves_completed)

simon |>
  mutate(
    simon_mobile_used = if_else(device_type == "Mobile", 1, 0),
    hour_submitted = hour(datetime)
  ) |>
  select(-datetime, -device_type)

simon |>
  group_by(pid, wave) |>
  summarise(
    has_na = any(if_any(ends_with("final"), is.na)),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = wave,
    values_from = has_na,
    names_prefix = "na_wave_"
  ) |>
  mutate(across(starts_with("na_wave_"), ~ replace_na(.x, TRUE))) |>
  count(na_wave_1, na_wave_3, na_wave_5)

simon_traits <- simon |>
  filter(!is.na(meanrt_final)) |>
  group_by(pid) |>
  summarise(
    simon_waves_count = n_distinct(wave),
    rt_mean = mean(meanrt_final, na.rm = TRUE),
    rt_sd = ifelse(n() > 1, sd(meanrt_final, na.rm = TRUE), 0),
    # roc: rate of change (gradient of rt over time)
    rt_roc = ifelse(
      n() >= 2,
      coef(lm(meanrt_final ~ wave))[2],
      0
    ),
    score_mean = mean(score_final, na.rm = TRUE),
    .groups = "drop"
  )

biweekly_mapper_data <- biweekly |>
  left_join(select(intake_participants, pid, adhd), by = "pid") |>
  filter(!is.na(adhd)) |>
  mutate(adhd = as.factor(adhd)) |>
  select(pid, adhd, wave, date, affective_valence, life_sat,
    starts_with("bangs_"), -starts_with("bangs_dup"), -bangs_failed_att_check,
    starts_with("gaming_value"),
    starts_with("wemwbs"),
    played_any_games
  ) |>
  # cuts a lot of bangs missingness
  filter(played_any_games == "Yes") |>
  drop_na() |>
  select(-played_any_games) |>
  left_join(biweekly_mapper_telemetry, by = join_by(pid, wave)) |>
  # add reaction time data
  left_join(simon_traits, by = "pid") |>
  drop_na()

summary(biweekly_mapper_data)
cat("Number of unique participants:", length(unique(biweekly_mapper_data$pid)))
cat("Dimension of mapper data:", dim(biweekly_mapper_data))
```

### Filter Functions

Due to the large amount of data,
we will need filter functions capable of explaining/understanding properties of ADHD participants,
and measuring the fundamental topology.

#### ADHD Explainability
The first lens/filter needs to try and represent the underlying properties that can describe ADHD participants.
I want to try either factor analysis of 3 compnents (for the ADHD groups) or some machine learning.
This methodology is inspired by the [KeplerMapper breast cancer example](https://kepler-mapper.scikit-tda.org/en/latest/generated/gallery/plot_breast_cancer.html).

##### PCA
```{r}
n_components <- 2

# data scaling
biweekly_mapper_data_scaled <- biweekly_mapper_data |>
  select(-pid, -date, -adhd, -binge_day_of_week) |>
  # scale by dividing by standard deviation
  mutate(across(everything(), ~scale(.)[,1]))

biweekly_mapper_pca <- prcomp(biweekly_mapper_data_scaled, center = FALSE, scale = FALSE)
summary(biweekly_mapper_pca)
```

1 or 2 principal components does not cover enough of the data's variance (0.16/0.33).
```{r}
ggplot(
  biweekly_mapper_pca$x,
  aes(x = PC1, y = PC2, colour = biweekly_mapper_data$adhd)
) +
  geom_point()

ggplot(biweekly_mapper_pca$x, aes(y = PC1)) +
  geom_histogram()
```

I do not think there is enough variance accounted for using PCA.

##### t-SNE
This is another dimensionality reduction method that converts the distances between points into probabilities.
The interest in using this method is because it is non-linear, which may overcome PCA's problem.

```{r}
set.seed(30)

# the scaled data was automatically centred
tsne_results <- Rtsne(
  biweekly_mapper_data_scaled,
  dims = 1,
  pca_center = FALSE,
  pca_scale = FALSE
)

tsne_vec <- as.vector(tsne_results$Y)
```

```{r}
(
  ggplot(as.data.frame(tsne_results$Y), aes(y = V1)) +
  geom_histogram()
) +
(
  ggplot(as.data.frame(tsne_results$Y), aes(y = V1)) +
  geom_boxplot()
)
```

Quick check to use the original data and let t-SNE handle the scaling.
```{r}
tsne_2 <- Rtsne(
  biweekly_mapper_data,
  dims = 2,
  # theta = 0.0, # exact t-SNE
  pca_center = TRUE,
  pca_scale = TRUE
)

ggplot(as.data.frame(tsne_2$Y), aes(x = V1, y = V2)) +
  geom_point(aes(colour = biweekly_mapper_data$adhd))
```

##### Decision Trees
My first idea is to create a decision tree capable of classifying ADHD participants.
From this, we can use the decision process as a filter (like the PCA principal components).
Decision trees are prone to overfitting, but in the case of exploring high-dimensional data, this can be okay.
We can limit the trees to stop this, or use a bagging method.

As we have imbalanced data, we need to change how we create the decision tree;
using accuracy as the evaluation metric will result in a model trying to just predict the majority class.

```{r}
biweekly_mapper_data_dt <- biweekly_mapper_data |>
  select(-pid, date)

# prediction recipe
adhd_recipe <- recipe(adhd ~ ., data = biweekly_mapper_data_dt)

# cross-validation for hyperparameter tuning
set.seed(30)
five_folds_cv <- vfold_cv(biweekly_mapper_data_dt, v = 5, strata = adhd)

# metrics focussing on imbalanced classes
imbal_metrics <- metric_set(pr_auc, bal_accuracy, sens, spec)

adhd_dt_spec <- decision_tree(
  cost_complexity = tune(),
  tree_depth = tune(),
  min_n = tune()
) |>
  set_engine(
    "rpart",
    # equal priors
    parms = list(prior = c(1/3, 1/3, 1/3))
  ) |>
  set_mode("classification")
adhd_dt_spec |> translate()

adhd_dt_wf <- workflow() |>
  add_recipe(adhd_recipe) |>
  add_model(adhd_dt_spec)

adhd_dt_grid <- grid_regular(
  cost_complexity(range = c(-8, -3)),
  tree_depth(range = c(4, 8)),
  min_n(range = c(2, 50)),
  levels = 5
)

# tuning decision tree
adhd_dt_results <- tune_grid(
  adhd_dt_wf,
  resamples = five_folds_cv,
  grid = adhd_dt_grid,
  metrics = imbal_metrics,
)

adhd_dt_best_params <- select_best(adhd_dt_results, metric = "pr_auc")

adhd_dt_final_wf <- finalize_workflow(adhd_dt_wf, adhd_dt_best_params)
adhd_dt_fit <- fit(adhd_dt_final_wf, data = biweekly_mapper_data_dt)

adhd_dt_best_params

adhd_dt <- extract_fit_engine(adhd_dt_fit)
summary(adhd_dt)

full_preds <- predict(
  adhd_dt_fit,
  new_data = biweekly_mapper_data_dt,
  type = "prob"
) |>
  bind_cols(biweekly_mapper_data_dt |> select(adhd))

adhd_dt_adhd_probs <- full_preds |>
  select(pred_diagnosed_adhd = `.pred_Diagnosed ADHD`) |>
  pull(pred_diagnosed_adhd)

adhd_threshold_perf <- full_preds |>
  # 2-level factor for classification
  mutate(
    adhd_binary = factor(
      if_else(adhd == "Diagnosed ADHD", "Diagnosed ADHD", "Other"),
      levels = c("Diagnosed ADHD", "Other")
    )
  ) |>
  # threshold_perf using the exact column name with backticks
  threshold_perf(
    truth = adhd_binary,
    estimate = `.pred_Diagnosed ADHD`,
    thresholds = seq(0.05, 0.95, 0.05)
  )
```

```{r}
rpart.plot(
  adhd_dt,
  type = 4, # split labels along the left/right branches
  extra = 104, # predicted class, class probabities, sample %
  under = TRUE,
  tweak = 1.1,
  box.palette = "BlGnYl"
)

ggplot(adhd_threshold_perf, aes(x = .threshold, y = .estimate, color = .metric)) +
  geom_line(linewidth = 1) +  # Updated size to linewidth for current ggplot2 versions
  labs(
    title = "Exploratory Threshold Performance for Diagnosed ADHD",
    x = "Probability Threshold", 
    y = "Score",
    color = "Metric"
  ) +
  theme_minimal()

optimal_threshold <- 0.3

multiclass_preds <- full_preds |>
  mutate(
    # winning class based on priority and probabilities
    .pred_class = case_when(
      `.pred_Diagnosed ADHD` >= optimal_threshold ~ "Diagnosed ADHD",
      `.pred_Neurotypical` >= `.pred_Self-Identified ADHD` ~ "Neurotypical",
      TRUE ~ "Self-Identified ADHD"
    ),
    .pred_class = factor(.pred_class, levels = levels(adhd))
  )

# confusion matrix
cm_3class <- multiclass_preds |>
  conf_mat(truth = adhd, estimate = .pred_class)

print(cm_3class)

# confusion matrix as heatmap
autoplot(cm_3class, type = "heatmap") +
  labs(
    title = "3-Class Confusion Matrix (Adjusted Threshold)",
    subtitle = paste("Diagnosed ADHD threshold set at", optimal_threshold),
    x = "Truth (Actual Class)",
    y = "Prediction (Model Output)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(face = "bold")
  )
```
This is okay, but not particularly great probabilities (they're fixed).

##### Boosted Decision Trees (XGBoost)

I will try a Boosted Decision Tree (XGBoost):
```{r}
set.seed(30)

xgb_data <- biweekly_mapper_data_scaled |>
  add_column(biweekly_mapper_data$adhd) |>
  rename(adhd = `biweekly_mapper_data$adhd`)

data_split <- initial_split(
  xgb_data,
  prop = 0.80,
  strata = adhd
)
train_data <- training(data_split)
test_data  <- testing(data_split)

# 5-fold cross validation for hyperparameter tuning
cv_folds <- vfold_cv(train_data, v = 5, strata = adhd)

# note: XGBoost requires all categorical predictors to be dummified
adhd_xgb_recipe <- recipe(adhd ~ ., data = train_data) |>
  # stop pid from being used in the the training
  # update_role(pid, new_role = "ID") |>
  # update_role(wave, new_role = "ID") |>
  # encode categorical 
  step_novel(all_nominal_predictors()) |>
  step_dummy(all_nominal_predictors(), one_hot = TRUE) |>
  # remove zero-variance features
  step_zv(all_predictors())
  # # optional (toggle)
  # themis::step_downsample(adhd, under_ratio = 1.5)

# boosted tree with hyperparameter tuning
adhd_xgb_spec <- boost_tree(
  trees = tune(), # number of boosting iterations (nrounds)
  tree_depth = tune(), # max depth of each individual tree
  min_n = tune(), # minimum child weight to stop splitting
  loss_reduction = tune(), # minimum gain to split further (gamma)
  sample_size = tune(), # subsample ratio of training instances
  mtry = tune(), # fraction of features sampled per tree
  learn_rate = tune() # learning rate / shrinkage (eta)
) |>
  set_engine("xgboost", objective = "multi:softprob") |>
  set_mode("classification")

# workflow
adhd_xgb_wf <- workflow() |>
  add_recipe(adhd_xgb_recipe) |>
  add_model(adhd_xgb_spec)

# hyperparameter Tuning Grid
# using Space-Filling Design to efficiently sample the parameter space
adhd_xgb_grid <- grid_space_filling(
  trees(range = c(50, 500)),
  tree_depth(range = c(3, 8)),
  min_n(range = c(2, 15)),
  loss_reduction(range = c(1e-8, 10), trans = log10_trans()),
  sample_size = sample_prop(range = c(0.5, 0.9)),
  finalize(mtry(), train_data),
  learn_rate(range = c(0.01, 0.2)),
  size = 25  # Evaluates 25 distinct hyperparameter combinations
)

# grid search on folds
cat("Starting hyperparameter tuning...\n")
adhd_xgb_res <- tune_grid(
  adhd_xgb_wf,
  resamples = cv_folds,
  grid = adhd_xgb_grid,
  metrics = metric_set(roc_auc, mn_log_loss, accuracy),
  control = control_grid(save_pred = TRUE, verbose = FALSE)
)

# choosing the best hyperparameters
# ROC AUC multi-class metric evaluates class separation across thresholds
best_params <- select_best(adhd_xgb_res, metric = "roc_auc")

adhd_final_xgb_wf <- finalize_workflow(adhd_xgb_wf, best_params)

# Fit on full training set and evaluate on test set
adhd_final_fit <- last_fit(adhd_final_xgb_wf, data_split)
collect_metrics(adhd_final_fit)

# Now I can extract the continuous lens, the probabilities:
# Fit final workflow on entire dataset
fitted_xgb <- fit(adhd_final_xgb_wf, data = xgb_data)

# Generate continuous class probability columns
adhd_xgb_probs <- augment(fitted_xgb, new_data = xgb_data) |>
  select(
    adhd,
    pred_diagnosed_adhd = `.pred_Diagnosed ADHD`,
    pred_neurotypical = `.pred_Neurotypical`,
    pred_self_idenfitied_adhd = `.pred_Self-Identified ADHD`
  )

# Inspect the smooth continuous probabilities ready for Mapper
head(adhd_xgb_probs)

# plot the probabilities of the classes, given the true class if diagnosed ADHD
adhd_xgb_probs |>
  pivot_longer(
    cols = starts_with("pred_"),
    names_to = "class_pred",
    values_to = "prob"
  ) |>
  filter(adhd == "Diagnosed ADHD") |>
  ggplot(aes(x = prob, fill = class_pred)) +
  geom_density(alpha = 0.5)

predictions <- predict(
  fitted_xgb,
  new_data = biweekly_mapper_data_scaled,
  type = "class"
) |>
  add_column(biweekly_mapper_data |> pull(adhd)) |>
  rename(
    truth = `pull(biweekly_mapper_data, adhd)`,
    pred = .pred_class
  )

xgb_cm <- predictions |>
  conf_mat(truth, pred)
```

```{r}
autoplot(xgb_cm, type = "heatmap")
```
This looks a lot better, and I will use the probability as the lens.


##### Isolation Forest
This classifies outliers in data, which could be useful in the ADHD data.

```{r}
set.seed(30)

biweekly_mapper_data_noadhd <- biweekly_mapper_data |>
  select(-pid, -adhd, -date)

is_outlier <- biweekly_mapper_data$adhd %in% c("Diagnosed ADHD")

model_orig <- isolation.forest(
  biweekly_mapper_data_noadhd,
  ndim = 1,
  sample_size = 256,
  ntrees = 500,
  missing_action = "fail"
)
pred_orig <- predict(model_orig, biweekly_mapper_data_noadhd)

model_dens <- isolation.forest(
  biweekly_mapper_data_noadhd,
  ndim = 1,
  sample_size = 256,
  ntrees = 500,
  missing_action = "fail",
  scoring_metric = "density"
)
pred_dens <- predict(model_dens, biweekly_mapper_data_noadhd)

model_fcf <- isolation.forest(
  biweekly_mapper_data_noadhd,
  ndim = 1,
  sample_size = 32,
  prob_pick_pooled_gain = 1,
  ntrees = 500,
  missing_action = "fail"
)
pred_fcf <- predict(model_fcf, biweekly_mapper_data_noadhd)

results_df <- data.frame(
    Model = c(
        "Isolation Forest",
        "Density Isolation Forest",
        "Fair-Cut Forest"
    ),
    AUROC = c(
        AUC(pred_orig, is_outlier),
        AUC(pred_dens, is_outlier),
        AUC(pred_fcf, is_outlier)
    )
)
results_df
```

#### Measures of Closeness

First, let us recap the structure of mapper data:
```{r}
summary(biweekly_mapper_data)
```

##### Distance Mean
Taken from the paper at doi.org/10.1109/jsen.2021.3130570.
This first calculates the mean of every column, the euclidean distances for each row and column,
then sum the column distances for each row.
This will produce a m x 1 matrix where m is the number of rows, which follows the same standard as
the previous filter functions.

```{r}
# these should be zero as the data is scaled
scaled_col_means <- colMeans(biweekly_mapper_data_scaled)

# the euclidean distance between two single points is just the absolute difference
A_dist <- sweep(biweekly_mapper_data_scaled, 2, scaled_col_means, "-")
dist_mean <- rowSums(A_dist^2)
```

```{r}
ggplot(as.data.frame(dist_mean), aes(x = dist_mean)) + geom_histogram()
```
The model appears to missclassify self-identified ADHD participants are Neurotypical rather than Diagnosed ADHD.

##### Distance from Neurotypical Average
I want to try calculating the column averages for the Neurotypical, then applying a similar method to dist mean.
```{r}
neurotypical_means <- biweekly_mapper_data_scaled |>
  add_column(biweekly_mapper_data$adhd, .before = TRUE) |>
  rename(adhd = `biweekly_mapper_data$adhd`) |>
  filter(adhd == "Neurotypical") |>
  select(-adhd) |>
  colMeans()

neuro_diffs <- sweep(biweekly_mapper_data_scaled, 2, neurotypical_means, "-")
neuro_dists <- rowSums(neuro_diffs^2)
```

##### Density Estimation
```{r}
mapper_dmatrix <- dist(biweekly_mapper_data_scaled)
mapper_density <- density_estimation(mapper_dmatrix)
density_lens <- as.vector(mapper_density$values)
```

```{r}
ggplot(as.data.frame(density_lens), aes(x = density_lens)) +
  geom_histogram()
```

##### L2 Norm
```{r}
l2_norms <- biweekly_mapper_data |>
  select(where(is.numeric)) |>
  apply(1, norm, type = "2")
```

#### Columns from the data
A quick look at some of the columns:
```{r}
ggplot(biweekly_mapper_data) +
  geom_point(aes(
    x = affective_valence,
    y = biweekly_mapper_pca$x[, 1],
  )) +
  facet_wrap(vars(adhd))

wemwbs_total <- biweekly_mapper_data |>
  mutate(
    wemwbs_total = wemwbs_1 +
      wemwbs_2 + wemwbs_3 + wemwbs_4 +
      wemwbs_5 + wemwbs_6 + wemwbs_7,
  ) |>
  pull(wemwbs_total) |>
  scale()

ggplot(biweekly_mapper_data) +
  geom_point(aes(
    x = wemwbs_total,
    y = wave,
  )) +
  facet_wrap(vars(adhd))
```

#### Evaluating Functions
Let's check the correlations:
```{r}
biweekly_mapper_filters <- tibble(
  adhd = biweekly_mapper_data$adhd,
  pc1 = biweekly_mapper_pca$x[, 1],
  pc2 = biweekly_mapper_pca$x[, 2],
  dt = adhd_dt_adhd_probs,
  xgb_adhd_pred = adhd_xgb_probs$pred_diagnosed_adhd,
  iso_forest = pred_orig,
  iso_forest_den = pred_dens,
  iso_forest_fc = pred_fcf,
  tsne = tsne_vec,
  dist_mean = dist_mean,
  density_estimated = density_lens,
  neuro_dist = neuro_dists,
  affective_val = biweekly_mapper_data$affective_valence,
  wemwbs_total = wemwbs_total[,1],
  l2_norm = l2_norms
)
filters_cor <- cor(select(biweekly_mapper_filters, -adhd))
filters_cor[] <- vapply(filters_cor, round, double(1), digits=2)
filters_cor
```

##### Spread Visualisations

```{r}
ggpairs(
  biweekly_mapper_filters,
  columns = 2:ncol(biweekly_mapper_filters),
  aes(colour = adhd, alpha = 0.5)
)
```
Across all filter combinations, the data is very dense.
This leads me to choose DBSCAN as the clustering algorithm.

## Mapper Trials
```{r}
make_mapper_filename <- function(
  mapper_name = "mapper",
  cover_type = NULL,
  intervals = NULL,
  overlap = NULL,
  width = NULL,
  method = NULL,
  method_params = NULL,
  ext = "json"
) {
  parts <- c()

  if (!is.null(mapper_name) && nzchar(mapper_name)) {
    parts <- c(parts, mapper_name)
  } else {
    parts <- c(parts, "mapper")
  }
  if (!is.null(cover_type) && nzchar(cover_type)) {
    parts <- c(parts, paste0("cov", cover_type))
  }
  if (!is.null(intervals) && nzchar(intervals)) {
    parts <- c(parts, paste0("int", intervals))
  }
  if (!is.null(overlap) && !is.na(overlap)) {
    parts <- c(parts, paste0("ov", overlap))
  }
  if (!is.null(width) && !is.na(width)) {
    parts <- c(parts, paste0("w", width))
  }
  if (!is.null(method) && !is.na(method) && nzchar(method)) {
    parts <- c(parts, method)
  }
  if (!is.null(method_params) && length(method_params) > 0) {
    valid_params <- method_params[!sapply(method_params, function(x) is.null(x) || is.na(x))]
    if (length(valid_params) > 0) {
      param_names <- names(valid_params)
      if (is.null(param_names)) param_names <- rep("", length(valid_params))
      param_str <- paste(param_names, unlist(valid_params), sep = "", collapse = "_")
      parts <- c(parts, param_str)
    }
  }
  ext <- sub("^\\.", "", ext)
  filename <- paste0(paste(parts, collapse = "_"), ".", ext)
  return(filename)
}

# add a colourisation feature
biweekly_mapper_data <- biweekly_mapper_data |>
  mutate(is_adhd = if_else(adhd == "Diagnosed ADHD", 1, 0)) |>
  relocate(adhd)
```
One of the filter functions should try to disperse the data, if possible by ADHD.
Examining the visualisations indicates that `dist_mean` or `PC1` is most capable at spreading the data.
PCA is designed to specifically extract the directions of maximum variance in the data.
However, it is not easily interpretable as it reduces the dimensions.
`dist_mean` is easily explainble as the sum of the residuals per row.
Therefore, I will choose `dist_mean`.

For the other filter, now that we have a metric that tries to disperse the data,
should aim to provide some explanation to the question I am exploring:
ADHD traits/features.
To do so requires some way of classifying the data by likelihood of having ADHD.

### First two principal components
```{r}
pca_lens <- biweekly_mapper_filters |>
  select(pc1, pc2) |>
  as.matrix()


# find bins to use (we have a large outlier for distances)
# biweekly_dists <- dist(biweekly_mapper_data_scaled)
# ggplot(as.data.frame(biweekly_dists), aes(x = x)) +
#   geom_histogram(bins = 80)

# mapper_pca_grid <- GridSearch(
#   original_data = biweekly_mapper_data_scaled,
#   filter_values = pca_lens,
#   label = biweekly_mapper_data$adhd,
#   column = "ADHD",
#   cover_type = "stride",
#   width_vec = c(0.2, 0.4, 0.5, 0.75, 1.0),
#   overlap_vec = c(20, 30, 50),
#   num_cores = 12,
#   out_dir = "examples/open-play/biweekly_pca_grid"
# )
```

```{r}
# mapper_name <- "biweekly_pca"
# cover_type <- "stride"
# intervals <- NULL
# overlap <- 50
# width <- 5
# method <- "dbscan"
# method_params <- list(eps = 5, minPts = 20)

mapper_name <- "biweekly_pca"
cover_type <- "stride"
intervals <- 10
overlap <- 30
width <- NULL
method <- "kmeans"
method_params <- list(max_kmeans_clusters = 2)

filename <- make_mapper_filename(
  mapper_name = mapper_name,
  cover_type = cover_type,
  intervals = intervals,
  overlap = overlap,
  width = width,
  method = method,
  method_params = method_params
)

cat("Creating file:", filename, "\n", sep = "")

pca_mapper <- MapperAlgo(
  original_data = biweekly_mapper_data_scaled,
  filter_values = pca_lens,
  percent_overlap = overlap,
  methods = method,
  method_params = method_params,
  cover_type = cover_type,
  intervals = intervals,
  interval_width = width,
  num_cores = 12
)

export_data <- list(
  adjacency = pca_mapper$adjacency,
  num_vertices = pca_mapper$num_vertices,
  level_of_vertex = pca_mapper$level_of_vertex,
  points_in_vertex = pca_mapper$points_in_vertex,
  original_data = biweekly_mapper_data
)

mapper_trials_dir <- file.path("examples", "open-play", "mappers")
if (!dir.exists(mapper_trials_dir)) {
  dir.create(mapper_trials_dir)
}

file.create(file.path(mapper_trials_dir, filename))
write(
  toJSON(export_data, auto_unbox = TRUE),
  file = here::here(file.path(mapper_trials_dir, filename))
)
# MapperPlotter(
#   pca_mapper,
#   original_data = biweekly_mapper_data_scaled,
#   label = biweekly_mapper_data$adhd
# )
```
Tried:
- DBSCAN(eps=0.3, minPts=5), 50%, width=0.4: massive graph too many nodes
- DBSCAN(eps=0.3, minPts=5), 30%, width=0.4: too many clusters to check, no structure
- DBSCAN(eps=0.3, minPts=20), 50%, width=0.4: improving, but still lots of central nodes with not many ADHD in
- Same with width = 1.5 and minPts = 40:
- Kmeans (3 clusters for ADHD), 10 intervals: too small nodes
- Same with 5 intervals, too many connections and overlap
- Now 30% overlap:
- Kmeans 2: for two ADHD groups
Increasing width to 5 gave nicer results, only 58 nodes, but we have too many people in some nodes (2000, a third), eps is 30?!

### Density Estimation & Dist Mean
```{r}
dist_dens_lens <- biweekly_mapper_filters |>
  select(dist_mean, density_estimated) |>
  as.matrix()

n_cols <- ncol(biweekly_mapper_data)

# minPts: >= d + 1, 2 x d
# eps: kNN distance graph to find elbow
k_val <- ncol(biweekly_mapper_data) + 1
kNNdistplot(biweekly_mapper_data_num, k = k_val)

dist_dens_mapper <- MapperAlgo(
  original_data = biweekly_mapper_data_num,
  filter_values = dist_dens_lens,
  percent_overlap = 50,
  methods = "dbscan",
  method_params = list(eps = 1.5, minPts = k_val),
  intervals = 10,
  num_cores = 10
)

dist_dens_mapper_single <- MapperAlgo(
  original_data = biweekly_mapper_data_num,
  filter_values = dist_dens_lens,
  percent_overlap = 50,
  methods = "hierarchical",
  method_params = list(method = "single", num_bins_when_clustering = 10),
  intervals = 10,
  num_cores = 10
)
```

### More meaningful columns

```{r} 
meaningful_lens <- biweekly_mapper_filters |>
  select(density_estimated, affective_val) |>
  add_column(biweekly_mapper_data$wave) |>
  rename(wave = `biweekly_mapper_data$wave`)
```

```{r}
mapper_name <- "biweekly_dens_affval_wave"
cover_type <- "stride"
intervals <- 10
overlap <- 50
width <- NULL
method <- "kmeans"
method_params <- list(max_kmeans_clusters = 2)

filename <- make_mapper_filename(
  mapper_name = mapper_name,
  cover_type = cover_type,
  intervals = intervals,
  overlap = overlap,
  width = width,
  method = method,
  method_params = method_params
)

cat("Creating file:", filename, "\n", sep = "")

meaningful_mapper <- MapperAlgo(
  original_data = biweekly_mapper_data_scaled,
  filter_values = meaningful_lens,
  percent_overlap = overlap,
  methods = method,
  method_params = method_params,
  cover_type = cover_type,
  intervals = intervals,
  interval_width = width,
  num_cores = 12
)

export_data <- list(
  adjacency = meaningful_mapper$adjacency,
  num_vertices = meaningful_mapper$num_vertices,
  level_of_vertex = meaningful_mapper$level_of_vertex,
  points_in_vertex = meaningful_mapper$points_in_vertex,
  original_data = biweekly_mapper_data
)

mapper_trials_dir <- file.path("examples", "open-play", "mappers")
if (!dir.exists(mapper_trials_dir)) {
  dir.create(mapper_trials_dir)
}

file.create(file.path(mapper_trials_dir, filename))
write(
  toJSON(export_data, auto_unbox = TRUE),
  file = here::here(file.path(mapper_trials_dir, filename))
)
```
